---
title: "Evaluation of Taxonomic Assignment Packages V4"
author: "Evan Krell, Chris Bird, Martin French"
date: "October 11, 2017"
output: html_document
runtime: shiny
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Libraries
library ("pracma")  # For string manipulation functions
library("stringr")
library(VennDiagram)
library("ggplot2")
#source("http://bioconductor.org/biocLite.R"); biocLite(c("graph", "RBGL"))
#install.packages("reshape")
#install.packages("Vennerable", repos="http://R-Forge.R-project.org")
library("Vennerable")
library("gridExtra")
```

## Load and prepare data

```{r load, echo = FALSE}

## SAP complete

#########################################################
# Load data common to all taxonomic assignment software #
#########################################################

setwd("~/Documents/code/genomics/taxonomic_evalutation/data")

###
# OTUs_seqs_samples: 
# 
# Every row has a unique sequence, the corrosponding OTU, 
# the ID of the source sample, and the count generated during a PCR-error corrective
# clustering step. Not to be confused with the count from clustering OTUs with CROP,
# whose count includes the PCR-error cluster count. 
###

OTUs_seqs_file <- "Simons.combo.nonchimeras.clean.OTU.samples_counts.BEST.csv"
OTUs_seqs_samples <- read.table (file = OTUs_seqs_file, header = FALSE, sep = ',', stringsAsFactors = FALSE)
colnames (OTUs_seqs_samples) <- c ("OTU_SEQID", "QSEQID", "SAMPLE_ID", "COUNT")

###
# OTUs_seqs: 
#
# A simplified form of OTUs_seqs_samples
# Simply maps sequences to their corrosponding OTUs
###


OTUs_seqs <- OTUs_seqs_samples[, c ("OTU_SEQID", "QSEQID")]
colnames (OTUs_seqs) <- c ("OTU_SEQID", "QSEQID")

###
# OTUs
#
# Simply a list of all OTUs
###

OTUs <- OTUs_seqs[!duplicated (OTUs_seqs$OTU_SEQID),]
OTUs <- as.list (OTUs$OTU_SEQID)

###
# Sample_IDs
#
# A list of Sample IDs
###

samples_file <- "Simons.samples.txt"
samples <- unlist (as.list (read.table (file = samples_file, header = FALSE, stringsAsFactors = FALSE)))
samples <- make.names (samples)


###
# Predators_Samples
#
# Table with samples and the predator associated.
###

pred_file <- "Predators2.csv"
Predators_Samples <- read.table (file = pred_file, header = TRUE, sep = ',', stringsAsFactors = FALSE)

##################################################
# Load data resulting from taxonomic assignment  #
##################################################

setwd("~/Documents/code/genomics/taxonomic_evalutation/newdata")



###
# ovt_XXX: 
#
# OTUs VS Tubes file. 
# Each row is an OTU, the results of taxonomic assignment 
# and the read count at a particular sample ID
###

ovt_vsearch_file <- "Simons_NDB.OVT.ann.vsearch.csv"
ovt_ecotag_file <- "Simons_NDB.OVT.ann.ecotag.csv"
ovt_blast_file <- "Simons_NDB.OVT.ann.blast.csv"
ovt_sap_file <- "Simons_NDB.OVT.SAP-C.csv"
ovt_vsearch <- read.table (file = ovt_vsearch_file, header = TRUE, sep = ',', stringsAsFactors = FALSE)
ovt_ecotag <- read.table (file = ovt_ecotag_file, header = TRUE, sep = ',', stringsAsFactors = FALSE)
ovt_blast <- read.table (file = ovt_blast_file,  header = TRUE, sep = ',', stringsAsFactors = FALSE)
ovt_sap <- read.table (file = ovt_sap_file, header = TRUE, sep = ',', stringsAsFactors = FALSE)

# Fix colname inconsistency
ovt_vsearch$X <- NULL
colnames (ovt_vsearch)[1:3] <- c ("OTU_SEQID", "OTU_SEQUENCE", "REF_SEQID")
ovt_ecotag$X <- NULL
ovt_ecotag <- ovt_ecotag[,c(1,2,4,3,5:ncol(ovt_ecotag))]
colnames (ovt_ecotag)[1:3] <- c ("OTU_SEQID", "OTU_SEQUENCE", "REF_SEQID")
ovt_blast$X.1 <- NULL
colnames (ovt_blast)[1:3] <- c ("OTU_SEQID", "OTU_SEQUENCE", "REF_SEQID")
ovt_sap$OTU_SEQUENCE <- rep.int (x = NA, times = nrow (ovt_sap))
ovt_sap$REF_SEQID <- rep.int (x = NA, times = nrow (ovt_sap))
ovt_sap <- ovt_sap[,c(1,ncol(ovt_sap)-1, ncol(ovt_sap),2:ncol(ovt_sap)-1)]
ovt_sap$OTU_SEQID.1 <- NULL
ovt_sap$OTU_SEQUENCE.1 <- NULL
colnames (ovt_sap)[1:3] <- c ("OTU_SEQID", "OTU_SEQUENCE", "REF_SEQID")
colnames (ovt_blast)[1:3] <- c ("OTU_SEQID", "OTU_SEQUENCE", "REF_SEQID")



# Remove unassigned sequences from ecotag results
ovt_ecotag <- ovt_ecotag[ovt_ecotag$NCBI_TAXID != '',]

# Use to query just the samples in an ovt_XXX
sample_vsearch_idx <- match(samples, colnames(ovt_vsearch))
sample_vsearch_idx <- sort(sample_vsearch_idx)
sample_ecotag_idx <- match(samples, colnames(ovt_ecotag))
sample_ecotag_idx <- sort(sample_ecotag_idx)
sample_blast_idx <- match(samples, colnames(ovt_blast))
sample_blast_idx <- sort(sample_blast_idx)
sample_sap_idx <- match(samples, colnames(ovt_sap))
sample_sap_idx <- sort(sample_sap_idx)

###
# charon_XXX:
#
# The Charon file format is a specification for a CSV
# that organizes the results of taxonomic assignment from various
# programs into a common format. 
# Is used as input to the production of OTUs VS Tubes.
#
# Contains the OTU sequence ID, The reference sequence ID, the assigned scientific name, 
# the assigned NCBI taxonomy ID, the sample ID, and the PCR-error correction merge count 
###

charon_vsearch_file <- "Simons_NDB.charon.vsearch.csv"
charon_ecotag_file <- "Simons_NDB.charon.ecotag.csv"
charon_blast_file <- "Simons_NDB.charon.blast.csv"
charon_vsearch <- read.table (file = charon_vsearch_file, header = FALSE, sep = ',', stringsAsFactors = FALSE)
charon_ecotag <- read.table (file = charon_ecotag_file, header = FALSE, sep = ',', stringsAsFactors = FALSE)
charon_blast <- read.table (file = charon_blast_file, header = FALSE, sep = ',', stringsAsFactors = FALSE)
colnames (charon_vsearch) <- c ("QUERY_SEQID", "REFERENCE_SEQID", "SCINAME", "NCBI_TAXID", "SAMPLE_ID", "COUNT", "OTU_SEQID")
colnames (charon_ecotag) <- c ("QUERY_SEQID", "REFERENCE_SEQID", "SCINAME", "NCBI_TAXID", "SAMPLE_ID", "COUNT", "OTU_SEQID")
colnames (charon_blast) <- c ("QUERY_SEQID", "REFERENCE_SEQID", "SCINAME", "NCBI_TAXID", "SAMPLE_ID", "COUNT", "OTU_SEQID")

# Remove unassigned sequences from ecotag results
charon_ecotag <- charon_ecotag[charon_ecotag$NCBI_TAXID != '1',]

# Fix formatting of Reference Sequence ID: Only have GI
# Specifically, the Ecotag results cut off part of the entire reference sequence ID,
# but only the GI is required for a unique comparison
ovt_vsearch$REF_SEQID <- str_split_fixed (ovt_vsearch$REF_SEQID, "\\|", 4)[,2]
ovt_ecotag$REF_SEQID <- str_split_fixed (ovt_ecotag$REF_SEQID, "\\|", 4)[,2]
ovt_blast$REF_SEQID <- str_split_fixed (ovt_blast$REF_SEQID, "\\|", 4)[,2]
charon_vsearch$REFERENCE_SEQID <- str_split_fixed (charon_vsearch$REFERENCE_SEQID, "\\|", 4)[,2]
charon_ecotag$REFERENCE_SEQID <- str_split_fixed (charon_ecotag$REFERENCE_SEQID, "\\|", 4)[,2]
charon_blast$REFERENCE_SEQID <- str_split_fixed (charon_blast$REFERENCE_SEQID, "\\|", 4)[,2]


###
# otu_data_XXX:
#
# Where Charon has data for each original sequence,
# this file compressed that information into only the OTUs.
# Contains the OTU sequence ID, the reference sequence ID, 
# assigned scientific name, and assigned NCBI taxonomy ID
### 

otu_data_vsearch <- charon_vsearch[with(charon_vsearch, order(charon_vsearch$OTU_SEQID)), ]
otu_data_vsearch <- otu_data_vsearch[!duplicated (otu_data_vsearch$OTU_SEQID),]
otu_data_vsearch$QUERY_SEQID <- NULL
otu_data_vsearch$SAMPLE_ID <- NULL
otu_data_vsearch$COUNT <- NULL

otu_data_ecotag <- charon_ecotag[with(charon_ecotag, order(charon_ecotag$OTU_SEQID)), ]
otu_data_ecotag <- otu_data_ecotag[!duplicated (otu_data_ecotag$OTU_SEQID),]
otu_data_ecotag$QUERY_SEQID <- NULL
otu_data_ecotag$SAMPLE_ID <- NULL
otu_data_ecotag$COUNT <- NULL

otu_data_blast <- charon_blast[with(charon_blast, order(charon_blast$OTU_SEQID)), ]
otu_data_blast <- otu_data_blast[!duplicated (otu_data_blast$OTU_SEQID),]
otu_data_blast$QUERY_SEQID <- NULL
otu_data_blast$SAMPLE_ID <- NULL
otu_data_blast$COUNT <- NULL

```

## Apply ID >= .85 filter (phylum) and >= .97 filter (species) to BLAST and VSEARCH

**Warning:** Is is a problem that I change some species assignments to phylum assignments without updating the TAXID? 

```{r ID_filter }

taxonomicRanksOfInterest <- c ("species", "genus", "family", "order", "class", "phylum", "kingdom")

# Vsearch
ovt_vsearch <- ovt_vsearch[ovt_vsearch$VSEARCH_IDENTITY_NT >= 85,]
ovt_vsearch_species <- ovt_vsearch[ovt_vsearch$VSEARCH_IDENTITY_NT >= 97,]
ovt_vsearch_phylum <- ovt_vsearch[ovt_vsearch$VSEARCH_IDENTITY_NT < 97,]
ovt_vsearch_phylum$species = ""
ovt_vsearch_phylum$genus = ""
ovt_vsearch_phylum$family = ""
ovt_vsearch_phylum$order = ""
ovt_vsearch_phylum$class = ""
ovt_vsearch = rbind.data.frame(ovt_vsearch_species, ovt_vsearch_phylum)

vsearch_taxonomy <- ovt_vsearch[,taxonomicRanksOfInterest]
vsearch_best_match_scinames <- apply(vsearch_taxonomy, 1, function(x) head(x[x != ""], 1))

vsearch_best_match_scinames <- unlist (lapply(X = vsearch_best_match_scinames, 
                                      function (x) ifelse(length(x) == 0, "", x[1])))
vsearch_best_match_levels <- apply(vsearch_taxonomy, 1, function(x) names(head(x[x != ""], 1)))
vsearch_best_match_levels <- unlist (lapply(X = vsearch_best_match_levels, 
                                      function (x) ifelse(length(x) == 0, "", x[1])))
vsearch_best_match <- cbind.data.frame (ovt_vsearch$OTU_SEQID, 
                                        vsearch_best_match_scinames, vsearch_best_match_levels)
colnames (vsearch_best_match) <- c ("OTU_SEQID", "SCINAME", "LEVEL")


# BLAST
ovt_blast <- ovt_blast[ovt_blast$BLAST_IDENTITY_NT >= 85,]
ovt_blast_species <- ovt_blast[ovt_blast$BLAST_IDENTITY_NT >= 97,]
ovt_blast_phylum <- ovt_blast[ovt_blast$BLAST_IDENTITY_NT < 97,]
ovt_blast_phylum$species = ""
ovt_blast_phylum$genus = ""
ovt_blast_phylum$family = ""
ovt_blast_phylum$order = ""
ovt_blast_phylum$class = ""
ovt_blast = rbind.data.frame(ovt_blast_species, ovt_blast_phylum)

blast_taxonomy <- ovt_blast[,taxonomicRanksOfInterest]
blast_best_match_scinames <- apply(blast_taxonomy, 1, function(x) head(x[x != ""], 1))

blast_best_match_scinames <- unlist (lapply(X = blast_best_match_scinames, 
                                      function (x) ifelse(length(x) == 0, "", x[1])))
blast_best_match_levels <- apply(blast_taxonomy, 1, function(x) names(head(x[x != ""], 1)))
blast_best_match_levels <- unlist (lapply(X = blast_best_match_levels, 
                                      function (x) ifelse(length(x) == 0, "", x[1])))
blast_best_match <- cbind.data.frame (ovt_blast$OTU_SEQID, 
                                        blast_best_match_scinames, blast_best_match_levels)
colnames (blast_best_match) <- c ("OTU_SEQID", "SCINAME", "LEVEL")


# ECOTAG
ecotag_taxonomy <- ovt_ecotag[,taxonomicRanksOfInterest]
ecotag_best_match_scinames <- apply(ecotag_taxonomy, 1, function(x) head(x[x != ""], 1))

ecotag_best_match_scinames <- unlist (lapply(X = ecotag_best_match_scinames, 
                                      function (x) ifelse(length(x) == 0, "", x[1])))
ecotag_best_match_levels <- apply(ecotag_taxonomy, 1, function(x) names(head(x[x != ""], 1)))
ecotag_best_match_levels <- unlist (lapply(X = ecotag_best_match_levels, 
                                      function (x) ifelse(length(x) == 0, "", x[1])))
ecotag_best_match <- cbind.data.frame (ovt_ecotag$OTU_SEQID, 
                                        ecotag_best_match_scinames, ecotag_best_match_levels)
colnames (ecotag_best_match) <- c ("OTU_SEQID", "SCINAME", "LEVEL")


#SAP 

sap_taxonomy <- ovt_sap[,taxonomicRanksOfInterest]
sap_best_match_scinames <- apply(sap_taxonomy, 1, function(x) head(x[x != ""], 1))

sap_best_match_scinames <- unlist (lapply(X = sap_best_match_scinames, 
                                      function (x) ifelse(length(x) == 0, "", x[1])))
sap_best_match_levels <- apply(sap_taxonomy, 1, function(x) names(head(x[x != ""], 1)))
sap_best_match_levels <- unlist (lapply(X = sap_best_match_levels, 
                                      function (x) ifelse(length(x) == 0, "", x[1])))
sap_best_match <- cbind.data.frame (ovt_sap$OTU_SEQID, 
                                        sap_best_match_scinames, sap_best_match_levels)
colnames (sap_best_match) <- c ("OTU_SEQID", "SCINAME", "LEVEL")




taxon_level_counts_vsearch <- sapply (X = taxonomicRanksOfInterest, 
                                      FUN = function (x) length (which(vsearch_best_match$LEVEL == x)))
taxon_level_counts_ecotag <- sapply (X = taxonomicRanksOfInterest, 
                                     FUN = function (x) length (which(ecotag_best_match$LEVEL == x)))
taxon_level_counts_blast <- sapply (X = taxonomicRanksOfInterest, 
                                     FUN = function (x) length (which(blast_best_match$LEVEL == x)))
taxon_level_counts_sap <- sapply (X = taxonomicRanksOfInterest, 
                                     FUN = function (x) length (which(sap_best_match$LEVEL == x)))

empty_taxon_count_vsearch <- length (which (vsearch_best_match$LEVEL == ""))
empty_taxon_count_ecotag <- length (which (ecotag_best_match$LEVEL == ""))
empty_taxon_count_blast <- length (which (blast_best_match$LEVEL == ""))
empty_taxon_count_sap <- length (which (sap_best_match$LEVEL == ""))
empty_counts <- c (empty_taxon_count_vsearch, empty_taxon_count_ecotag, empty_taxon_count_blast, empty_taxon_count_sap)


total_count_vsearch <- sum (taxon_level_counts_vsearch) + empty_taxon_count_vsearch
total_count_ecotag <- sum (taxon_level_counts_ecotag) + empty_taxon_count_ecotag
total_count_blast <- sum (taxon_level_counts_blast) + empty_taxon_count_blast
total_count_sap <- sum (taxon_level_counts_sap) + empty_taxon_count_sap


total_counts <- c (total_count_vsearch, total_count_ecotag, total_count_blast, total_count_sap)
   

# ASSERTION: total_counts should equal the nrow (ovt_XXX_85)
    
titles_col <- c ("Vsearch", "Ecotag", "Blast", "SAP")
counts_table <- rbind.data.frame (taxon_level_counts_vsearch, taxon_level_counts_ecotag, taxon_level_counts_blast, taxon_level_counts_sap)
counts_table <- cbind.data.frame(titles_col, counts_table, empty_counts, total_counts)
colnames (counts_table) <- c ("Source", taxonomicRanksOfInterest, "unassigned", "total")

percent_assigned <- counts_table$species / total_counts

print ("Counts assigned to species level, with an id >= .85 filter for Blast and Vsearch")
print (counts_table)

print ("Percent assigned to species level, with an id >= .85 filter for Blast and Vsearch")
print (percent_assigned)

```




## Filter singletons

In order to reduce amount of junk assignments, remove based on the following criteria:
  
1. Singleton (type 1): The read count across all samples is 1
2. Singleton (type 2): The read count in any single sample is never > 1
3. Singleton (type 3): The read count in any single sample is never greater than 1% of total count across all samples

``` {r filter 1}

## SAP Complete

## Filter singleton (type 1)
# 
#singleton_1_vsearch_length <- nrow (ovt_vsearch[ovt_vsearch$TOTAL == 1,])
# ovt_vsearch <- ovt_vsearch[ovt_vsearch$TOTAL > 1,]
# 
#singleton_1_ecotag_length <- nrow (ovt_ecotag[ovt_ecotag$TOTAL == 1,])
# ovt_ecotag <- ovt_ecotag[ovt_ecotag$TOTAL > 1,]
# 
#singleton_1_blast_length <- nrow (ovt_blast[ovt_blast$TOTAL == 1,])
# ovt_blast <- ovt_blast[ovt_blast$TOTAL > 1,]
# 
#singleton_1_sap_length <- nrow (ovt_sap[ovt_sap$TOTAL == 1,])
# ovt_sap <- ovt_sap[ovt_sap$TOTAL > 1,]
# 
# ## Filter singleton (type 2)
# 
#singleton_2_vsearch_length <- nrow (ovt_vsearch[apply(ovt_vsearch[,sample_vsearch_idx], 1, function(row) {all(row <= 1)}),])
# ovt_vsearch <- ovt_vsearch[apply(ovt_vsearch[,sample_vsearch_idx], 1, function(row) {any(row > 1)}),]
# 
#singleton_2_ecotag_length <- nrow (ovt_ecotag[apply(ovt_ecotag[,sample_ecotag_idx], 1, function(row) {all(row <= 1)}),])
# ovt_ecotag <- ovt_ecotag[apply(ovt_ecotag[,sample_ecotag_idx], 1, function(row) {any(row > 1)}),]
# 
#singleton_2_blast_length <- nrow (ovt_blast[apply(ovt_blast[,sample_blast_idx], 1, function(row) {all(row <= 1)}),])
# ovt_blast <- ovt_blast[apply(ovt_blast[,sample_blast_idx], 1, function(row) {any(row > 1)}),]
# 
#singleton_2_sap_length <- nrow (ovt_sap[apply(ovt_sap[,sample_sap_idx], 1, function(row) {all(row <= 1)}),])
# ovt_sap <- ovt_sap[apply(ovt_sap[,sample_sap_idx], 1, function(row) {any(row > 1)}),]
# 
# ## Filter singleton (type 3)
# 
#singleton_3_vsearch_length <- nrow (ovt_vsearch [apply(ovt_vsearch[,sample_vsearch_idx], 1, function(row) {  all( .01 >= (row/sum(row) ) ) }), ])
# ovt_vsearch <- ovt_vsearch [apply(ovt_vsearch[,sample_vsearch_idx], 1, function(row) {  any( .01 < (row/sum(row) ) ) }), ]
# 
#singleton_3_ecotag_length <- nrow (ovt_ecotag [apply(ovt_ecotag[,sample_ecotag_idx], 1, function(row) {  all( .01 >= (row/sum(row) ) ) }), ])
# ovt_ecotag <- ovt_ecotag [apply(ovt_ecotag[,sample_ecotag_idx], 1, function(row) {  any( .01 < (row/sum(row) ) ) }), ]
# 
#singleton_3_blast_length <- nrow (ovt_blast [apply(ovt_blast[,sample_blast_idx], 1, function(row) {  all( .01 >= (row/sum(row) ) ) }), ])
# ovt_blast <- ovt_blast [apply(ovt_blast[,sample_blast_idx], 1, function(row) {  any( .01 < (row/sum(row) ) ) }), ]
# 
#singleton_3_sap_length <- nrow (ovt_sap [apply(ovt_sap[,sample_sap_idx], 1, function(row) {  all( .01 >= (row/sum(row) ) ) }), ])
# ovt_sap <- ovt_sap [apply(ovt_sap[,sample_sap_idx], 1, function(row) {  any( .01 < (row/sum(row) ) ) }), ]
# 
# # Total the filtered OTUs
#total_filtered_vsearch <- singleton_1_vsearch_length + singleton_2_vsearch_length + singleton_3_vsearch_length
#total_filtered_ecotag <- singleton_1_ecotag_length + singleton_2_ecotag_length + singleton_3_ecotag_length
#total_filtered_blast <- singleton_1_blast_length + singleton_2_blast_length + singleton_3_blast_length
#total_filtered_sap <- singleton_1_sap_length + singleton_2_sap_length + singleton_3_sap_length

######|######| GOES BELOW THIS EXECUTION BOX
################| | Vsearch | Ecotag  | Blast | SAP
################| - |:-:| :-:| :-: | -:
################  | Singletons (type 1) | `r singleton_1_vsearch_length` | `r singleton_1_ecotag_length` | `r singleton_1_blast_length` | `r singleton_1_sap_length`
################| Singletons (type 2) | `r singleton_2_vsearch_length` | `r singleton_2_ecotag_length` | `r singleton_2_blast_length` | `r singleton_2_sap_length`
################| Singletons (type 3) | `r singleton_3_vsearch_length` | `r singleton_3_ecotag_length` | `r singleton_3_blast_length` | `r singleton_3_sap_length`
################| Total Filtered OTUs| `r total_filtered_vsearch` | `r total_filtered_ecotag` | `r total_filtered_blast` |  `r total_filtered_sap`
################| Remaining OTUs     | `r nrow (ovt_vsearch)` | `r nrow (ovt_ecotag)` | `r nrow (ovt_blast)` | `r nrow (ovt_sap)`


```




## Overview

```{r compare, echo = FALSE}

## Not SAP compatable.

total_OTUs <- length (OTUs)

uniq_vsearch_tseqs <- unique(ovt_vsearch$REF_SEQID)
uniq_ecotag_tseqs <- unique(ovt_ecotag$REF_SEQID)
uniq_blast_tseqs <- unique(ovt_blast$REF_SEQID)
uniq_sap_tseqs <- NA

uniq_vsearch_tseqs_len <- length (uniq_vsearch_tseqs)
uniq_ecotag_tseqs_len <- length (uniq_ecotag_tseqs)
uniq_blast_tseqs_len <- length (uniq_blast_tseqs)
uniq_sap_tseqs_len <- NA

intersect_len_EV <- length (intersect (uniq_vsearch_tseqs, uniq_ecotag_tseqs))
intersect_len_EB <- length (intersect (uniq_blast_tseqs, uniq_ecotag_tseqs))
intersect_len_VB <- length (intersect (uniq_vsearch_tseqs, uniq_blast_tseqs))
inter <- intersect (uniq_vsearch_tseqs, uniq_ecotag_tseqs)
intersect_len <- length (intersect (inter, uniq_blast_tseqs))

un <- union(uniq_vsearch_tseqs, uniq_ecotag_tseqs)
union_len <- length (union (un, uniq_blast_tseqs))

vn <- list (0)
vn[[1]] <- uniq_vsearch_tseqs
vn[[2]] <- uniq_ecotag_tseqs
vn[[3]] <- uniq_blast_tseqs
names (vn) <- c ("Vsearch", "Ecotag", "Blast")

```

The following table compares the number of OTUs present in the data
produced for each method of taxonomic assignment.

In addition, the table shows the number of unique reference assignments for each
assignment method, as well as the intersection of all three: 
  the number of OTUs that were assigned to the same reference sequence. 

| | Vsearch | Ecotag  | Blast | SAP
| - |:-:| :-:| :-: | -:
  | Number of Assigned OTUs | `r nrow (ovt_vsearch)` | `r nrow (ovt_ecotag)` | `r nrow (ovt_blast)` | `r length (unique (ovt_sap$OTU_SEQID))`
| Total OTUs | `r total_OTUs` | `r total_OTUs` | `r total_OTUs` | `r total_OTUs`
| Percent Assigned OTUs | `r (nrow (ovt_vsearch)/total_OTUs)*100` | `r (nrow (ovt_ecotag)/total_OTUs)*100` | `r (nrow (ovt_blast)/total_OTUs)*100` | `r ((length (unique (ovt_sap$OTU_SEQID)))/total_OTUs)*100`
| Unique Reference Sequences      | `r uniq_vsearch_tseqs_len` | `r uniq_ecotag_tseqs_len` | `r uniq_blast_tseqs_len` | `r uniq_sap_tseqs_len`
| Intersection of Unique Reference Sequences        | `r intersect_len`      |   `r intersect_len` | `r intersect_len` | `r NA`
| Percent References Sequences Shared | `r (intersect_len/uniq_vsearch_tseqs_len)*100`      |    `r (intersect_len/uniq_ecotag_tseqs_len)*100` | `r (intersect_len/uniq_blast_tseqs_len)*100` | `r NA`


## Venn diagram of OTUs assigned to same reference

```{r venn, echo = FALSE}

## Not SAP compatable

grid.newpage()

plot (Venn(vn)) # For numbers
plot (Venn(vn), show = list (SetLabels = FALSE, FaceText = "")) # For presentation
```

## Venn diagram showing intersection of assigned OTUs

```{r venn_assigned_OTUs, echo = FALSE}

intersect_len_EV <- length (intersect (ovt_ecotag$OTU_SEQID, ovt_vsearch$OTU_SEQID))
interset_len_EB <- length (intersect (ovt_ecotag$OTU_SEQID, ovt_blast$OTU_SEQID))
intersect_len_VB <- length (intersect (ovt_blast$OTU_SEQID, ovt_vsearch$OTU_SEQID))

intersect_len_ES <- length (intersect (ovt_ecotag$OTU_SEQID, ovt_sap$OTU_SEQID))
intersect_len_VS <- length (intersect (ovt_vsearch$OTU_SEQID, ovt_sap$OTU_SEQID))
intersect_len_BS <- length (intersect (ovt_blast$OTU_SEQID, ovt_sap$OTU_SEQID))

inter <- intersect (ovt_ecotag$OTU_SEQID, ovt_vsearch$OTU_SEQID)
intersect_len_EVB <- length (intersect (inter, ovt_blast$OTU_SEQID))
inter <- intersect (ovt_ecotag$OTU_SEQID, ovt_vsearch$OTU_SEQID)
intersect_len_EVS <- length (intersect (inter, ovt_sap$OTU_SEQID))
inter <- intersect (ovt_blast$OTU_SEQID, ovt_vsearch$OTU_SEQID)
intersect_len_BVS <- length (intersect (inter, ovt_sap$OTU_SEQID))


inter <- intersect (ovt_ecotag$OTU_SEQID, ovt_vsearch$OTU_SEQID)
inter <- intersect (inter, ovt_blast$OTU_SEQID)
inter <- intersect (inter, ovt_sap$OTU_SEQID)
intersect_len_EVBS <- length (inter)

vn_otu <- list (0)
vn_otu[[1]] <- ovt_vsearch$OTU_SEQID
vn_otu[[2]] <- ovt_ecotag$OTU_SEQID
vn_otu[[3]] <- ovt_blast$OTU_SEQID
vn_otu[[4]] <- ovt_sap$OTU_SEQID

names (vn_otu) <- c ("Vsearch", "Ecotag", "Blast", "SAP")

# The "fudged" version is to have the ecotag and blast portions actually visible
#vn_otu_fudged <- list (0)
#vn_otu_fudged[[1]] <- ovt_vsearch$OTU_SEQID
#vn_otu_fudged[[2]] <- unlist ( c(ovt_ecotag$OTU_SEQID, as.character (c(1:50))))
#vn_otu_fudged[[3]] <- unlist ( c(ovt_blast$OTU_SEQID, as.character (c(50:100))))
#vn_otu_fudged[[4]] <- ovt_sap$OTU_SEQID
#names (vn_otu_fudged) <- c ("Vsearch", "Ecotag", "Blast", "SAP")

grid.newpage()

plot (Venn (vn_otu), type = "squares", doWeights = FALSE) # For numbers
#plot (Venn (vn_otu_fudged), show = list (SetLabels = FALSE, FaceText = "")) # For presentation

```


| Intersection | Number of Assigned OTUs |
|  | - | -:
|  Ecotag, Vsearch | `r intersect_len_EV`
|Ecotag, Blast | `r intersect_len_EB`
|Ecotag, SAP | `r intersect_len_VS`
|Vsearch, Blast | `r intersect_len_VB`
|Vsearch, SAP | `r intersect_len_VS`
|Blast, SAP | `r intersect_len_BS`
|Ecotag, Vsearch, Blast | `r intersect_len_EVB`
|Ecotag, Vsearch, SAP | `r intersect_len_EVS`
|Vsearch, Blast, SAP | `r intersect_len_BVS`
|Ecotag, Vsearch, Blast, SAP | `r intersect_len_EVBS`

## Investigate phylogenetic levels of taxonomic assingment of OTUs

The following table shows, for each phylogenetic level,
the number of OTU assignments where that level was the lowest assignment. 
This table may also reveal issues where there is not lowest. 
This could be some kind of error with the assignment, or a behavior of the
assignment package. For example, ecotag keeps all data in the results. 
The sequences that it cannot assign, it assigns to the "root" of the phylogenetic tree.


```{r phylogeny}

## SAP complete

taxonomicRanksOfInterest <- c ("species", "genus", "family", "order", "class", "phylum", "kingdom")

# Vsearch
vsearch_taxonomy <- ovt_vsearch[,taxonomicRanksOfInterest]
vsearch_best_match_scinames <- apply(vsearch_taxonomy, 1, function(x) head(x[x != ""], 1))

vsearch_best_match_scinames <- unlist (lapply(X = vsearch_best_match_scinames, 
                                              function (x) ifelse(length(x) == 0, "", x[1])))
vsearch_best_match_levels <- apply(vsearch_taxonomy, 1, function(x) names(head(x[x != ""], 1)))
vsearch_best_match_levels <- unlist (lapply(X = vsearch_best_match_levels, 
                                            function (x) ifelse(length(x) == 0, "", x[1])))
vsearch_best_match <- cbind.data.frame (ovt_vsearch$OTU_SEQID, 
                                        vsearch_best_match_scinames, vsearch_best_match_levels)
colnames (vsearch_best_match) <- c ("OTU_SEQID", "SCINAME", "LEVEL")


# Ecotag
ecotag_taxonomy <- ovt_ecotag[,taxonomicRanksOfInterest]
ecotag_best_match_scinames <- apply(ecotag_taxonomy, 1, function(x) head(x[x != ""], 1))

ecotag_best_match_scinames <- unlist (lapply(X = ecotag_best_match_scinames, 
                                             function (x) ifelse(length(x) == 0, "", x[1])))
ecotag_best_match_levels <- apply(ecotag_taxonomy, 1, function(x) names(head(x[x != ""], 1)))
ecotag_best_match_levels <- unlist (lapply(X = ecotag_best_match_levels, 
                                           function (x) ifelse(length(x) == 0, "", x[1])))
ecotag_best_match <- cbind.data.frame (ovt_ecotag$OTU_SEQID, 
                                       ecotag_best_match_scinames, ecotag_best_match_levels, ovt_ecotag$ECOTAG_IDENTITY_NT)
colnames (ecotag_best_match) <- c ("OTU_SEQID", "SCINAME", "LEVEL", "ID")


# Blast
blast_taxonomy <- ovt_blast[,taxonomicRanksOfInterest]
blast_best_match_scinames <- apply(blast_taxonomy, 1, function(x) head(x[x != ""], 1))

blast_best_match_scinames <- unlist (lapply(X = blast_best_match_scinames, 
                                            function (x) ifelse(length(x) == 0, "", x[1])))
blast_best_match_levels <- apply(blast_taxonomy, 1, function(x) names(head(x[x != ""], 1)))
blast_best_match_levels <- unlist (lapply(X = blast_best_match_levels, 
                                          function (x) ifelse(length(x) == 0, "", x[1])))
blast_best_match <- cbind.data.frame (ovt_blast$OTU_SEQID, 
                                      blast_best_match_scinames, blast_best_match_levels)
colnames (blast_best_match) <- c ("OTU_SEQID", "SCINAME", "LEVEL")


# SAP
sap_taxonomy <- ovt_sap[,taxonomicRanksOfInterest]
sap_best_match_scinames <- apply(sap_taxonomy, 1, function(x) head(x[x != ""], 1))

sap_best_match_scinames <- unlist (lapply(X = sap_best_match_scinames, 
                                          function (x) ifelse(length(x) == 0, "", x[1])))
sap_best_match_levels <- apply(sap_taxonomy, 1, function(x) names(head(x[x != ""], 1)))
sap_best_match_levels <- unlist (lapply(X = sap_best_match_levels, 
                                        function (x) ifelse(length(x) == 0, "", x[1])))
sap_best_match <- cbind.data.frame (ovt_sap$OTU_SEQID, 
                                    sap_best_match_scinames, sap_best_match_levels)
colnames (sap_best_match) <- c ("OTU_SEQID", "SCINAME", "LEVEL")



taxon_level_counts_vsearch <- sapply (X = taxonomicRanksOfInterest, 
                                      FUN = function (x) length (which(vsearch_best_match$LEVEL == x)))
taxon_level_counts_ecotag <- sapply (X = taxonomicRanksOfInterest, 
                                     FUN = function (x) length (which(ecotag_best_match$LEVEL == x)))
taxon_level_counts_blast <- sapply (X = taxonomicRanksOfInterest, 
                                    FUN = function (x) length (which(blast_best_match$LEVEL == x)))
taxon_level_counts_sap <- sapply (X = taxonomicRanksOfInterest, 
                                  FUN = function (x) length (which(sap_best_match$LEVEL == x)))


empty_taxon_count_vsearch <- length (which (vsearch_best_match$LEVEL == ""))
empty_taxon_count_ecotag <- length (which (ecotag_best_match$LEVEL == ""))
empty_taxon_count_blast <- length (which (blast_best_match$LEVEL == ""))
empty_taxon_count_sap <- length (which (sap_best_match$LEVEL == ""))
empty_counts <- c (empty_taxon_count_vsearch, empty_taxon_count_ecotag, empty_taxon_count_blast, empty_taxon_count_sap)


total_count_vsearch <- sum (taxon_level_counts_vsearch) + empty_taxon_count_vsearch
total_count_ecotag <- sum (taxon_level_counts_ecotag) + empty_taxon_count_ecotag
total_count_blast <- sum (taxon_level_counts_blast) + empty_taxon_count_blast
total_count_sap <- sum (taxon_level_counts_sap) + empty_taxon_count_sap
total_counts <- c (total_count_vsearch, total_count_ecotag, total_count_blast, total_count_sap)


# ASSERTION: total_counts should equal the nrow (ovt_XXX)

titles_col <- c ("Vsearch", "Ecotag", "Blast", "SAP")
counts_table <- rbind.data.frame (taxon_level_counts_vsearch, taxon_level_counts_ecotag, taxon_level_counts_blast, taxon_level_counts_sap)
counts_table <- cbind.data.frame(titles_col, counts_table, empty_counts, total_counts)
colnames (counts_table) <- c ("Source", taxonomicRanksOfInterest, "unranked", "total")

percent_assigned <- counts_table$species / total_counts

print (counts_table)

print ("Percent assigned to species level")
print (percent_assigned)

# bar plot of proportion assigned at species level
df <- cbind.data.frame (counts_table$Source, percent_assigned)
colnames (df) <- c ("source", "proportions")
ggplot (data=df, aes(y=proportions, x=source, fill=source)) + geom_bar(stat="identity", color = "black" ) + scale_fill_manual(values=c("#bebada", "#8dd3c7", "#ffaabc", "#ffffb3" ))+ theme_bw()


# Explore ranks of Ecotag best matches
ecotag_rank_id <- ecotag_best_match[,3:4]
sapply (X = ecotag_rank_id$LEVEL, FUN = function (x){ if (!(x %in% taxonomicRanksOfInterest) ){"unranked"}else{x}})
query <- unlist (c (taxonomicRanksOfInterest )) #, "unranked"))

# bar plot rank assignments for Ecotag  
proportions <- unlist (counts_table[2,][2:8][1,] / total_counts[[2]])
df <- data.frame(taxon_rank = query, proportions)
ggplot (data=df, aes(y=proportions, x=taxon_rank)) + geom_bar(stat="identity", fill = "#8dd3c7", color = "black") + scale_x_discrete (limits = query) + scale_y_continuous(labels = scales::percent)+ theme_bw()

ggplot (data=ecotag_rank_id, aes(y=ID, x=LEVEL)) + geom_boxplot(fill = "#8dd3c7", colour = "black" ) + scale_x_discrete (limits = query)+ theme_bw() + ylab ("Identity Score") + xlab ("Taxonomic Rank")

# Explore ranks of SAP best matches
# bar plot rank assignments for SAP
query <- unlist (c (taxonomicRanksOfInterest )) #, "unranked"))
proportions <- unlist (counts_table[4,][2:8][1,] / total_counts[[2]])
df <- data.frame(taxon_rank = query, proportions)
ggplot (data=df, aes(y=proportions, x=taxon_rank)) + geom_bar(stat="identity", fill = "#8dd3c7", color = "black") + scale_x_discrete (limits = query) + scale_y_continuous(labels = scales::percent)+ theme_bw()

# Commented out because it used the ID score (which SAP does not have)
#vsearch_rank_id <- vsearch_best_match[,c('SCINAME', 'LEVEL')]
#sapply (X = vsearch_rank_id$LEVEL, FUN = function (x){ if (!(x %in% taxonomicRanksOfInterest) ){"unranked"}else{x}})
#ggplot (data=vsearch_rank_id, aes(y=ID, x=LEVEL)) + geom_boxplot(fill = "#8dd3c7", colour = "black" ) + scale_x_discrete (limits = query)




```



## Investigate identity scores

BLAST, Vsearch and Ecotag each return an identity score.

The Vsearch parameters specified a minimum identity score of 0.7.

The Ecotag parameters specified a minumum identity score of 0.7.

The BLAST parameters specified a minimum identity score of 0. (????)

The following is intended to see if the assignment methods give similar identity scores when assigned to the same reference sequence,
as well as the overall spread of identity scores. 

```{r identity_scores}

# Ignore SAP

## Get each assigned OTU with identity score
otu_identity_vsearch <- cbind.data.frame (ovt_vsearch$OTU_SEQID,
                                          ovt_vsearch$VSEARCH_IDENTITY_NT / 100, ovt_vsearch$REF_SEQID)
colnames (otu_identity_vsearch) <- c ("OTU_SEQID", "VSEARCH_IDENTITY_NT", "VSEARCH_REF_SEQID")


otu_identity_ecotag <- cbind.data.frame (ovt_ecotag$OTU_SEQID, 
                                         ovt_ecotag$ECOTAG_IDENTITY_NT, ovt_ecotag$REF_SEQID)
colnames (otu_identity_ecotag) <- c ("OTU_SEQID", "ECOTAG_IDENTITY_NT", "ECOTAG_REF_SEQID")


otu_identity_blast <- cbind.data.frame (ovt_blast$OTU_SEQID,
                                        ovt_blast$BLAST_IDENTITY_NT / 100, ovt_blast$REF_SEQID)
colnames (otu_identity_blast) <- c ("OTU_SEQID", "BLAST_IDENTITY_NT", "BLAST_REF_SEQID")


# Together such that each OTU has all identity scores
otu_identity_EV <- merge (x = otu_identity_vsearch, y = otu_identity_ecotag, by = "OTU_SEQID")
otu_identity_VB <- merge (x = otu_identity_vsearch, y = otu_identity_blast, by = "OTU_SEQID")
otu_identity_EB <- merge (x = otu_identity_ecotag, y = otu_identity_blast, by = "OTU_SEQID")
otu_identity_EVB <- merge (x = otu_identity_EV, y = otu_identity_blast,  by = "OTU_SEQID")


par(mfrow=c(1,3))
plot(x = otu_identity_EV$VSEARCH_IDENTITY_NT, y = otu_identity_EV$ECOTAG_IDENTITY_NT,
     ylab = "Ecotag Identity", xlab = "Vsearch Identity", col = "black", bg = "#80b1d3", pch=21)
#abline (lm (otu_identity_EV$VSEARCH_IDENTITY_NT ~ otu_identity_EV$ECOTAG_IDENTITY_NT))
text(.82, 1, paste0 ("n = ", nrow (otu_identity_EV)), cex = 2)

plot(x = otu_identity_VB$VSEARCH_IDENTITY_NT, y = otu_identity_VB$BLAST_IDENTITY_NT,
     ylab = "Blast Identity", xlab = "Vsearch Identity", col = "black", bg = "#fb8072", pch=21)
#abline (lm (otu_identity_VB$VSEARCH_IDENTITY_NT ~ otu_identity_VB$BLAST_IDENTITY_NT))
text(.82, 1, paste0 ("n = ", nrow (otu_identity_VB)), cex = 2)

plot(x = otu_identity_EB$BLAST_IDENTITY_NT, y = otu_identity_EB$ECOTAG_IDENTITY_NT,
     ylab = "Ecotag Identity", xlab = "Blast Identity", col = "black", bg = "#fdb462", pch=21)
#abline (lm (otu_identity_EB$BLAST_IDENTITY_NT ~ otu_identity_EB$ECOTAG_IDENTITY_NT))
text(.84, 1, paste0 ("n = ", nrow (otu_identity_EB)), cex = 2)
title("Identity Scores between assignments from same OTU", outer=TRUE, line = -3)

```


*2-sample t-test between Ecotag Identity and Vsearch Identity for same OTUs*
  ```{r}
var.test(x = otu_identity_EV$VSEARCH_IDENTITY_NT, y = otu_identity_EV$ECOTAG_IDENTITY_NT)
t.test(otu_identity_EV$VSEARCH_IDENTITY_NT, otu_identity_EV$ECOTAG_IDENTITY_NT, var.equal=FALSE, paired=TRUE, alternative = "two.sided")
```
*2-sample t-test between Blast Identity and Vsearch Identity for same OTUs*
  ```{r}
var.test(x = otu_identity_VB$VSEARCH_IDENTITY_NT, y = otu_identity_VB$BLAST_IDENTITY_NT)
t.test(otu_identity_VB$VSEARCH_IDENTITY_NT, otu_identity_VB$BLAST_IDENTITY_NT, var.equal=FALSE, paired=TRUE, alternative = "two.sided")
```
*2-sample t-test between Ecotag Identity and Ecotag Identity for same OTUs*
  ```{r}
var.test(x = otu_identity_EB$BLAST_IDENTITY_NT, y = otu_identity_EB$ECOTAG_IDENTITY_NT)
t.test(otu_identity_EB$BLAST_IDENTITY_NT, otu_identity_EB$ECOTAG_IDENTITY_NT, var.equal=FALSE, paired=TRUE, alternative = "two.sided")
```





-----
  
  ## How similar are the score for when two methods reach the same assignment for an OTU?
  
```{r }
otu_EV_same_refseq <- otu_identity_EV[as.character (otu_identity_EV$VSEARCH_REF_SEQID) == as.character (otu_identity_EV$ECOTAG_REF_SEQID),]
otu_EV_same_refseq <- otu_EV_same_refseq[complete.cases(otu_EV_same_refseq),]
num_same_ref_seq_EV <- nrow (otu_EV_same_refseq)

otu_VB_same_refseq <- otu_identity_VB[ as.character (otu_identity_VB$VSEARCH_REF_SEQID) == as.character (otu_identity_VB$BLAST_REF_SEQID),]
otu_VB_same_refseq <- otu_VB_same_refseq[complete.cases(otu_VB_same_refseq),]
num_same_ref_seq_VB <- nrow (otu_VB_same_refseq)

otu_EB_same_refseq <- otu_identity_EB[ as.character (otu_identity_EB$ECOTAG_REF_SEQID) == as.character (otu_identity_EB$BLAST_REF_SEQID),]
otu_EB_same_refseq <- otu_EB_same_refseq[complete.cases(otu_EB_same_refseq),]
num_same_ref_seq_EB <- nrow (otu_EB_same_refseq)

```

Vsearch and Ecotag share `r num_same_ref_seq_EV` assignments to the same reference sequence.

Vsearch and Blast share `r num_same_ref_seq_VB` assignments to the same reference sequence.

Ecotag and Blast share `r num_same_ref_seq_EB` assignments to the same reference sequence.

```{r}
par(mfrow=c(1,3))
plot(x = otu_EV_same_refseq$VSEARCH_IDENTITY_NT, y = otu_EV_same_refseq$ECOTAG_IDENTITY_NT,
     ylab = "Ecotag Identity", xlab = "Vsearch Identity", col = "black", bg = "#80b1d3", pch=21)
text(.86, 1, paste0 ("n = ", nrow (otu_EV_same_refseq)), cex = 2)

plot(x = otu_VB_same_refseq$VSEARCH_IDENTITY_NT, y = otu_VB_same_refseq$BLAST_IDENTITY_NT,
     ylab = "Blast Identity", xlab = "Vsearch Identity", col = "black", bg = "#fb8072", pch = 21)
text(.85, 1, paste0 ("n = ", nrow (otu_VB_same_refseq)), cex = 2)

plot(x = otu_EB_same_refseq$BLAST_IDENTITY_NT, y = otu_EB_same_refseq$ECOTAG_IDENTITY_NT,
     ylab = "Ecotag Identity", xlab = "Blast Identity", col = "black", bg = "#fdb462", pch = 21) 
text(.87, 1, paste0 ("n = ", nrow (otu_EB_same_refseq)), cex = 2)

title("Identity Scores when OTU is assigned to same reference sequence", outer=TRUE, line = -3)

```

*2-sample t-test between Ecotag Identity and Vsearch Identity for matched reference sequence*
```{r}
var.test(x = otu_EV_same_refseq$VSEARCH_IDENTITY_NT, y = otu_EV_same_refseq$ECOTAG_IDENTITY_NT)
t.test(otu_EV_same_refseq$VSEARCH_IDENTITY_NT, otu_EV_same_refseq$ECOTAG_IDENTITY_NT, var.equal=TRUE, paired=TRUE, alternative = "two.sided")
```
*2-sample t-test between Blast Identity and Vsearch Identity for matched reference sequence*
```{r}
var.test(x = otu_VB_same_refseq$VSEARCH_IDENTITY_NT, y = otu_VB_same_refseq$BLAST_IDENTITY_NT)
t.test(otu_VB_same_refseq$VSEARCH_IDENTITY_NT, otu_VB_same_refseq$BLAST_IDENTITY_NT , var.equal=FALSE, paired=TRUE, alternative = "two.sided")
```
*2-sample t-test between Ecotag Identity and Ecotag Identity for matched reference sequence*
```{r}
var.test(x = otu_EB_same_refseq$BLAST_IDENTITY_NT, y = otu_EB_same_refseq$ECOTAG_IDENTITY_NT)
t.test(otu_EB_same_refseq$BLAST_IDENTITY_NT , otu_EB_same_refseq$ECOTAG_IDENTITY_NT, var.equal=TRUE, paired=TRUE, alternative = "two.sided")
```


## Investigate Taxonomy


```{r taxonomy, echo = FALSE}

# SAP complete

# Count number of unique ranks per phylogenetic level

uniqRanks <- cbind ( c ("Vsearch", "Ecotag", "Blast", "SAP"), rbind (
  apply(vsearch_taxonomy, MARGIN = 2, FUN = function (x) length (unique(x))),
  apply(ecotag_taxonomy, MARGIN = 2, FUN = function (x) length (unique(x))),
  apply(blast_taxonomy, MARGIN = 2, FUN = function (x) length (unique(x))),
  apply(sap_taxonomy, MARGIN = 2, FUN = function (x) length (unique(x)))) )
colnames (uniqRanks)[1] <- "Source"
print (uniqRanks)

intersect_taxonomy <- function (A, B){
  intersect_lens <- c ()
  for (i in taxonomicRanksOfInterest){
    intersect_lens <- c (intersect_lens, ( length ( intersect ( unlist (unique (A[i])), unlist (unique (B[i])) ) )) )
  }
  intersect_lens
}

intersect_taxonomy_three <- function (A, B, C){
  intersect_lens <- c ()
  for (i in taxonomicRanksOfInterest){
    inter_AB <- intersect (unlist (unique (A[i])), unlist (unique (B[i])) )
    intersect_lens <- c (intersect_lens, ( length ( intersect ( inter_AB, unlist (unique (C[i])) ) )) )
  }
  intersect_lens
}

intersect_taxonomy_four <- function (A, B, C, D){
  intersect_lens <- c ()
  for (i in taxonomicRanksOfInterest){
    inter_AB <- intersect (unlist (unique (A[i])), unlist (unique (B[i])) )
    inter_ABC <- intersect (inter_AB, unlist (unique (C[i])) )
    intersect_lens <- c (intersect_lens, ( length ( intersect ( inter_ABC, unlist (unique (D[i])) ) )) )
  }
  intersect_lens
}

intersect_counts_EV <- intersect_taxonomy(A = vsearch_taxonomy, B = ecotag_taxonomy)
intersect_counts_EB <- intersect_taxonomy(A = ecotag_taxonomy, B = blast_taxonomy)
intersect_counts_BV <- intersect_taxonomy(A = vsearch_taxonomy, B = blast_taxonomy)

intersect_counts_ES <- intersect_taxonomy(A = ecotag_taxonomy, B = sap_taxonomy)
intersect_counts_VS <- intersect_taxonomy(A = vsearch_taxonomy, B = sap_taxonomy)
intersect_counts_BS <- intersect_taxonomy(A = blast_taxonomy, B = sap_taxonomy)

intersect_counts_EBV <- intersect_taxonomy_three(A = vsearch_taxonomy, B = ecotag_taxonomy, C = blast_taxonomy)
intersect_counts_EVS <- intersect_taxonomy_three(A = vsearch_taxonomy, B = ecotag_taxonomy, C = sap_taxonomy)
intersect_counts_BVS <- intersect_taxonomy_three(A = vsearch_taxonomy, B = sap_taxonomy, C = blast_taxonomy)

intersect_counts_EBVS <- intersect_taxonomy_four(A = vsearch_taxonomy, B = ecotag_taxonomy, C = blast_taxonomy, D = sap_taxonomy)


intersect_counts_table <- cbind.data.frame (intersect_counts_EV, intersect_counts_EB, intersect_counts_ES, 
                                            intersect_counts_BV, intersect_counts_VS, intersect_counts_BS,
                                            intersect_counts_EBV, intersect_counts_EVS, intersect_counts_BVS,
                                            intersect_counts_EBVS)
colnames (intersect_counts_table) <- c ("Ecotag, Vsearch", "Ecotag, Blast", "Ecotag, SAP",
                                        "Vsearch, Blast", "Vsearch, SAP", "Blast, SAP",
                                        "Ecotag, Vsearch, Blast", "Ecotag, Vsearch, SAP", "Vsearch, Blast, SAP",
                                        "Ecotag, Vsearch, Blast, SAP")
rownames (intersect_counts_table) <- taxonomicRanksOfInterest
print (intersect_counts_table)


uniqRanks <- cbind ( c ("Vsearch", "Ecotag", "Blast", "SAP"), rbind (
  apply(vsearch_taxonomy, MARGIN = 2, FUN = function (x) length (unique(x))),
  apply(ecotag_taxonomy, MARGIN = 2, FUN = function (x) length (unique(x))),
  apply(blast_taxonomy, MARGIN = 2, FUN = function (x) length (unique(x))),
  apply(sap_taxonomy, MARGIN = 2, FUN = function (x) length (unique(x))) ) )
colnames (uniqRanks)[1] <- "Source"

print (uniqRanks)

# Get counts of specific unique ranks within a level
vsearch_taxonomy_uniq <- apply(vsearch_taxonomy, MARGIN = 2, FUN = function (x) unique(x))
ecotag_taxonomy_uniq <- apply(ecotag_taxonomy, MARGIN = 2, FUN = function (x) unique(x))
blast_taxonomy_uniq <- apply(blast_taxonomy, MARGIN = 2, FUN = function (x) unique(x))
sap_taxonomy_uniq <- apply(sap_taxonomy, MARGIN = 2, FUN = function (x) unique(x))

# Vsearch class counts
vsearch_class_counts <- sapply (vsearch_taxonomy_uniq$class, FUN = function (x) length (which(vsearch_taxonomy$class == x)))
vsearch_phylum_counts <- sapply (vsearch_taxonomy_uniq$phylum, FUN = function (x) length (which(vsearch_taxonomy$phylum == x)))
vsearch_kingdom_counts <- sapply (vsearch_taxonomy_uniq$kingdom, FUN = function (x) length (which(vsearch_taxonomy$kingdom == x)))

# Ecotag class counts
ecotag_class_counts <- sapply (ecotag_taxonomy_uniq$class, FUN = function (x) length (which(ecotag_taxonomy$class == x)))
ecotag_phylum_counts <- sapply (ecotag_taxonomy_uniq$phylum, FUN = function (x) length (which(ecotag_taxonomy$phylum == x)))
ecotag_kingdom_counts <- sapply (ecotag_taxonomy_uniq$kingdom, FUN = function (x) length (which(ecotag_taxonomy$kingdom == x)))

# Blast class counts
blast_class_counts <- sapply (blast_taxonomy_uniq$class, FUN = function (x) length (which(blast_taxonomy$class == x)))
blast_phylum_counts <- sapply (blast_taxonomy_uniq$phylum, FUN = function (x) length (which(blast_taxonomy$phylum == x)))
blast_kingdom_counts <- sapply (blast_taxonomy_uniq$kingdom, FUN = function (x) length (which(blast_taxonomy$kingdom == x)))

# SAP class counts
sap_class_counts <- sapply (sap_taxonomy_uniq$class, FUN = function (x) length (which(sap_taxonomy$class == x)))
sap_phylum_counts <- sapply (sap_taxonomy_uniq$phylum, FUN = function (x) length (which(sap_taxonomy$phylum == x)))
sap_kingdom_counts <- sapply (sap_taxonomy_uniq$kingdom, FUN = function (x) length (which(sap_taxonomy$kingdom == x)))

print ("Unique class counts (no filter)")
print (vsearch_class_counts)
print (ecotag_class_counts)
print (blast_class_counts)
print (sap_class_counts)

print ("Unique phylum counts (no filter)")
print (vsearch_phylum_counts)
print (ecotag_phylum_counts)
print (blast_phylum_counts)
print (sap_phylum_counts)

print ("Unique kingdom counts (no filter)")
print (vsearch_kingdom_counts)
print (ecotag_kingdom_counts)
print (blast_kingdom_counts)
print (sap_kingdom_counts)

```


### Venn diagrams of unique phylum intersection

```{r taxon_venn_phylum}

# SAP complete

vn_phylum <- list (0)
vn_phylum[[1]] <- vsearch_taxonomy_uniq$phylum
vn_phylum[[2]] <- ecotag_taxonomy_uniq$phylum
vn_phylum[[3]] <- blast_taxonomy_uniq$phylum
vn_phylum[[4]] <- sap_taxonomy_uniq$phylum
names (vn_phylum) <- c ("Vsearch", "Ecotag", "Blast", "SAP")
plot (Venn (vn_phylum), doWeights = TRUE)


plot (Venn (vn_phylum), doWeights = TRUE, show = list (SetLabels = FALSE, FaceText = ""))

```


## Proportion of OTUs assigned to species

**Warning:** Is this not problematic because many OTUs are not even returned for SAP? 
  
```{r}

# SAP complete

# bar plot of proportion assigned at species level
df <- cbind.data.frame (counts_table$Source, percent_assigned)
colnames (df) <- c ("source", "proportions")
ggplot (data=df, aes(y=proportions, x=source, fill=source)) + geom_bar(stat="identity", color = "black" ) + scale_fill_manual(values=c("#bebada", "#8dd3c7", "#ffaabc", "#ffffb3" ))+ theme_bw()

# bar plot rank assignments for Ecotag  
proportions <- unlist (counts_table[2,][2:8][1,] / sum (counts_table[2,][2:8][1,] ))
df_97 <- cbind.data.frame (proportions, taxonomicRanksOfInterest)
df_97 <- data.frame(taxon_rank = taxonomicRanksOfInterest, proportions)

# bar plot rank assignments for SAP
query <- unlist (c (taxonomicRanksOfInterest )) #, "unranked"))
proportions <- unlist (counts_table[4,][2:8][1,] / sum (counts_table[4,][2:8][1,] ))
df <- data.frame(taxon_rank = query, proportions)
ggplot (data=df, aes(y=proportions, x=taxon_rank)) + geom_bar(stat="identity", fill = "#8dd3c7", color = "black") + scale_x_discrete (limits = query) + scale_y_continuous(labels = scales::percent)+ theme_bw() + ylab ("Assignment Proportion") + xlab ("Taxonomic Rank")

```

## Counts of uniques at various taxonomic levels

```{r taxon}

# SAP complete

uniqRanks <- cbind ( c ("Vsearch", "Ecotag", "Blast", "SAP"), rbind (
  apply(vsearch_taxonomy, MARGIN = 2, FUN = function (x) length (unique(x))),
  apply(ecotag_taxonomy, MARGIN = 2, FUN = function (x) length (unique(x))),
  apply(blast_taxonomy, MARGIN = 2, FUN = function (x) length (unique(x))),
  apply(sap_taxonomy, MARGIN = 2, FUN = function (x) length (unique(x)))) )
colnames (uniqRanks)[1] <- "Source"
print (uniqRanks)


intersect_counts_EV <- intersect_taxonomy(A = vsearch_taxonomy, B = ecotag_taxonomy)
intersect_counts_EB <- intersect_taxonomy(A = ecotag_taxonomy, B = blast_taxonomy)
intersect_counts_BV <- intersect_taxonomy(A = vsearch_taxonomy, B = blast_taxonomy)

intersect_counts_ES <- intersect_taxonomy(A = ecotag_taxonomy, B = sap_taxonomy)
intersect_counts_VS <- intersect_taxonomy(A = vsearch_taxonomy, B = sap_taxonomy)
intersect_counts_BS <- intersect_taxonomy(A = blast_taxonomy, B = sap_taxonomy)

intersect_counts_EBV <- intersect_taxonomy_three(A = vsearch_taxonomy, B = ecotag_taxonomy, C = blast_taxonomy)
intersect_counts_EVS <- intersect_taxonomy_three(A = vsearch_taxonomy, B = ecotag_taxonomy, C = sap_taxonomy)
intersect_counts_BVS <- intersect_taxonomy_three(A = vsearch_taxonomy, B = sap_taxonomy, C = blast_taxonomy)

intersect_counts_EBVS <- intersect_taxonomy_four(A = vsearch_taxonomy, B = ecotag_taxonomy, C = blast_taxonomy, D = sap_taxonomy)


intersect_counts_table <- cbind.data.frame (intersect_counts_EV, intersect_counts_EB, intersect_counts_ES, 
                                            intersect_counts_BV, intersect_counts_VS, intersect_counts_BS,
                                            intersect_counts_EBV, intersect_counts_EVS, intersect_counts_BVS,
                                            intersect_counts_EBVS)
colnames (intersect_counts_table) <- c ("Ecotag, Vsearch", "Ecotag, Blast", "Ecotag, SAP",
                                        "Vsearch, Blast", "Vsearch, SAP", "Blast, SAP",
                                        "Ecotag, Vsearch, Blast", "Ecotag, Vsearch, SAP", "Vsearch, Blast, SAP",
                                        "Ecotag, Vsearch, Blast, SAP")
rownames (intersect_counts_table) <- taxonomicRanksOfInterest
print (intersect_counts_table)


uniqRanks <- cbind ( c ("Vsearch", "Ecotag", "Blast", "SAP"), rbind (
  apply(vsearch_taxonomy, MARGIN = 2, FUN = function (x) length (unique(x))),
  apply(ecotag_taxonomy, MARGIN = 2, FUN = function (x) length (unique(x))),
  apply(blast_taxonomy, MARGIN = 2, FUN = function (x) length (unique(x))),
  apply(sap_taxonomy, MARGIN = 2, FUN = function (x) length (unique(x))) ) )
colnames (uniqRanks)[1] <- "Source"

print (uniqRanks)

# Get counts of specific unique ranks within a level
vsearch_taxonomy_uniq <- apply(vsearch_taxonomy, MARGIN = 2, FUN = function (x) unique(x))
ecotag_taxonomy_uniq <- apply(ecotag_taxonomy, MARGIN = 2, FUN = function (x) unique(x))
blast_taxonomy_uniq <- apply(blast_taxonomy, MARGIN = 2, FUN = function (x) unique(x))
sap_taxonomy_uniq <- apply(sap_taxonomy, MARGIN = 2, FUN = function (x) unique(x))

# Vsearch class counts
vsearch_class_counts <- sapply (vsearch_taxonomy_uniq$class, FUN = function (x) length (which(vsearch_taxonomy$class == x)))
vsearch_phylum_counts <- sapply (vsearch_taxonomy_uniq$phylum, FUN = function (x) length (which(vsearch_taxonomy$phylum == x)))
vsearch_kingdom_counts <- sapply (vsearch_taxonomy_uniq$kingdom, FUN = function (x) length (which(vsearch_taxonomy$kingdom == x)))

# Ecotag class counts
ecotag_class_counts <- sapply (ecotag_taxonomy_uniq$class, FUN = function (x) length (which(ecotag_taxonomy$class == x)))
ecotag_phylum_counts <- sapply (ecotag_taxonomy_uniq$phylum, FUN = function (x) length (which(ecotag_taxonomy$phylum == x)))
ecotag_kingdom_counts <- sapply (ecotag_taxonomy_uniq$kingdom, FUN = function (x) length (which(ecotag_taxonomy$kingdom == x)))

# Blast class counts
blast_class_counts <- sapply (blast_taxonomy_uniq$class, FUN = function (x) length (which(blast_taxonomy$class == x)))
blast_phylum_counts <- sapply (blast_taxonomy_uniq$phylum, FUN = function (x) length (which(blast_taxonomy$phylum == x)))
blast_kingdom_counts <- sapply (blast_taxonomy_uniq$kingdom, FUN = function (x) length (which(blast_taxonomy$kingdom == x)))

# SAP class counts
sap_class_counts <- sapply (sap_taxonomy_uniq$class, FUN = function (x) length (which(sap_taxonomy$class == x)))
sap_phylum_counts <- sapply (sap_taxonomy_uniq$phylum, FUN = function (x) length (which(sap_taxonomy$phylum == x)))
sap_kingdom_counts <- sapply (sap_taxonomy_uniq$kingdom, FUN = function (x) length (which(sap_taxonomy$kingdom == x)))

print ("Unique class counts")
print (vsearch_class_counts)
print (ecotag_class_counts)
print (blast_class_counts)
print (sap_class_counts)

print ("Unique phylum counts")
print (vsearch_phylum_counts)
print (ecotag_phylum_counts)
print (blast_phylum_counts)
print (sap_phylum_counts)

print ("Unique kingdom counts")
print (vsearch_kingdom_counts)
print (ecotag_kingdom_counts)
print (blast_kingdom_counts)
print (sap_kingdom_counts)

```

## Venn diagrams of uniques at various taxonomic levels

```{r taxon_venn}

par(mfrow=c(1,3))
vn_species <- list (0)
vn_species[[1]] <- vsearch_taxonomy_uniq$species
vn_species[[2]] <- ecotag_taxonomy_uniq$species
vn_species[[3]] <- blast_taxonomy_uniq$species
vn_species[[4]] <- sap_taxonomy$species
names (vn_species) <- c ("Vsearch", "Ecotag", "Blast", "SAP")
#plot (Venn (vn_species), doWeights = FALSE)
#plot (Venn (vn_species), doWeights = TRUE)
#plot (Venn (vn_species), doWeights = TRUE, show = list (SetLabels = FALSE, FaceText = ""))
plot (Venn (vn_species), type = "squares", doWeights = FALSE)


vn_genus <- list (0)
vn_genus[[1]] <- vsearch_taxonomy_uniq$genus
vn_genus[[2]] <- ecotag_taxonomy_uniq$genus
vn_genus[[3]] <- blast_taxonomy_uniq$genus
vn_genus[[4]] <- sap_taxonomy$genus
names (vn_genus) <- c ("Vsearch", "Ecotag", "Blast", "SAP")
#plot (Venn (vn_genus), doWeights = FALSE)
plot (Venn (vn_genus), type = "squares", doWeights = FALSE)


vn_family <- list (0)
vn_family[[1]] <- vsearch_taxonomy_uniq$family
vn_family[[2]] <- ecotag_taxonomy_uniq$family
vn_family[[3]] <- blast_taxonomy_uniq$family
vn_family[[4]] <- sap_taxonomy$family
names (vn_family) <- c ("Vsearch", "Ecotag", "Blast", "SAP")
#plot (Venn (vn_family), doWeights = FALSE)
plot (Venn (vn_family), type = "squares", doWeights = FALSE)


vn_order <- list (0)
vn_order[[1]] <- vsearch_taxonomy_uniq$order
vn_order[[2]] <- ecotag_taxonomy_uniq$order
vn_order[[3]] <- blast_taxonomy_uniq$order
vn_order[[4]] <- sap_taxonomy$order
names (vn_order) <- c ("Vsearch", "Ecotag", "Blast", "SAP")
#plot (Venn (vn_order), doWeights = FALSE)
plot (Venn (vn_order), type = "squares", doWeights = FALSE)


vn_class <- list (0)
vn_class[[1]] <- vsearch_taxonomy_uniq$class
vn_class[[2]] <- ecotag_taxonomy_uniq$class
vn_class[[3]] <- blast_taxonomy_uniq$class
vn_class[[4]] <- sap_taxonomy$class
names (vn_class) <- c ("Vsearch", "Ecotag", "Blast", "SAP")
#plot (Venn (vn_class), doWeights = FALSE)
plot (Venn (vn_class), type = "squares", doWeights = FALSE)


vn_phylum <- list (0)
vn_phylum[[1]] <- vsearch_taxonomy_uniq$phylum
vn_phylum[[2]] <- ecotag_taxonomy_uniq$phylum
vn_phylum[[3]] <- blast_taxonomy_uniq$phylum
vn_phylum[[4]] <- sap_taxonomy$phylum
names (vn_phylum) <- c ("Vsearch", "Ecotag", "Blast", "SAP")
#plot (Venn (vn_phylum), doWeights = TRUE)
plot (Venn (vn_phylum), type = "squares", doWeights = FALSE)


```


## Investigate preditors

```{r preditors}

# SAP todo

Predators_Samples$SAMPLE_ID <- make.names (Predators_Samples$SAMPLE_ID)


# Build critters vs tubes

##ovt_vsearch[ovt_vsearch$species == "Pseudorhombila quadridentata", "X558.A"]

# Fix a sample ID error
colnames (ovt_vsearch_species) <- gsub (x = colnames (ovt_vsearch_species), pattern = "B0", replacement = "B")
colnames (ovt_ecotag) <- gsub (x = colnames (ovt_ecotag), pattern = "B0", replacement = "B")
colnames (ovt_blast_species) <- gsub (x = colnames (ovt_blast), pattern = "B0", replacement = "B")
colnames (ovt_sap) <- gsub (x = colnames (ovt_sap), pattern = "B0", replacement = "B")

predator_match_vsearch <- apply (X = Predators_Samples, MARGIN = 1, FUN = function (x) { sum(ovt_vsearch_species[ovt_vsearch_species$species == x[2], x[1]])  })
predator_match_len_vsearch <- length (which (sapply (X = predator_match_vsearch, FUN = function (x) {x > 0})))
predator_match_proportion_vsearch <- predator_match_len_vsearch / length (samples)

predator_match_ecotag <- apply (X = Predators_Samples, MARGIN = 1, FUN = function (x) { sum (ovt_ecotag[ovt_ecotag$species == x[[2]], x[[1]]])  })
predator_match_len_ecotag <- length (which (sapply (X = predator_match_ecotag, FUN = function (x) {x > 0})))
predator_match_proportion_ecotag <- predator_match_len_ecotag / length (samples)

predator_match_blast <- apply (X = Predators_Samples, MARGIN = 1, FUN = function (x) { sum (ovt_blast_species[ovt_blast_species$species == x[[2]], x[[1]]])  })
predator_match_len_blast <- length (which (sapply (X = predator_match_blast, FUN = function (x) {x > 0})))
predator_match_proportion_blast <- predator_match_len_blast / length (samples)

predator_match_sap <- apply (X = Predators_Samples, MARGIN = 1, FUN = function (x) { sum (ovt_sap[ovt_sap$species == x[[2]], x[[1]]])  })
predator_match_len_sap <- length (which (sapply (X = predator_match_sap, FUN = function (x) {x > 0})))
predator_match_proportion_sap <- predator_match_len_sap / length (samples)

predator_match_proportions <- cbind.data.frame ( c ("Vsearch", "Ecotag", "Blast", "SAP"),
                                                 c (predator_match_proportion_vsearch, predator_match_proportion_ecotag, 
                                                    predator_match_proportion_blast, predator_match_proportion_sap))
colnames(predator_match_proportions) <- c ("method", "proportions")


p1 <- ggplot (data=predator_match_proportions, aes(y=proportions, x=method, fill=method)) + geom_bar(stat="identity", color = "black" ) + scale_fill_manual(values=c("#bebada", "#8dd3c7", "#ffaabc", "#ffffb3")) + ylim(0,1) + theme_bw()

plot (p1)
```

```{r predators_unique_venn}

blast_taxonomy_95 <- ovt_blast_species[,taxonomicRanksOfInterest]
vsearch_taxonomy_95 <- ovt_vsearch_species[,taxonomicRanksOfInterest]

# Get counts of specific unique ranks within a level
vsearch_taxonomy_uniq <- apply(vsearch_taxonomy_95, MARGIN = 2, FUN = function (x) unique(x))
ecotag_taxonomy_uniq <- apply(ecotag_taxonomy, MARGIN = 2, FUN = function (x) unique(x))
blast_taxonomy_uniq <- apply(blast_taxonomy_95, MARGIN = 2, FUN = function (x) unique(x))
sap_taxonomy_uniq <- apply(sap_taxonomy, MARGIN = 2, FUN = function (x) unique(x))






uniqPredators <- unique (Predators_Samples$SCINAME)
uniqPredatorsGenus <- unique (Predators_Samples$GENUS)

vn_preds <- list (0)
vn_preds[[1]] <- intersect (vsearch_taxonomy_uniq$species, uniqPredators)
vn_preds[[2]] <- intersect (ecotag_taxonomy_uniq$species, uniqPredators)
vn_preds[[3]] <- intersect (blast_taxonomy_uniq$species, uniqPredators)
vn_preds[[4]] <- intersect (sap_taxonomy_uniq$species, uniqPredators)
names (vn_preds) <- c ("Vsearch", "Ecotag", "Blast", "SAP")
#plot (Venn (vn_phylum), doWeights = TRUE)
plot (Venn (vn_preds), type = "squares", doWeights = FALSE)

vn_predsGenus <- list (0)
vn_predsGenus[[1]] <- intersect (vsearch_taxonomy_uniq$genus, uniqPredatorsGenus)
vn_predsGenus[[2]] <- intersect (ecotag_taxonomy_uniq$genus, uniqPredatorsGenus)
vn_predsGenus[[3]] <- intersect (blast_taxonomy_uniq$genus, uniqPredatorsGenus)
vn_predsGenus[[4]] <- intersect (sap_taxonomy_uniq$genus, uniqPredatorsGenus)
names (vn_predsGenus) <- c ("Vsearch", "Ecotag", "Blast", "SAP")
#plot (Venn (vn_phylum), doWeights = TRUE)
plot (Venn (vn_predsGenus), type = "squares", doWeights = FALSE)



```

```{r predators_unique_bar}
predator_unique_counts <- cbind.data.frame ( c ("Vsearch", "Ecotag", "Blast", "SAP"),
                                                 c (length (vn_preds[[1]]), length (vn_preds[[2]]), 
                                                    length (vn_preds[[3]]), length (vn_preds[[4]])))
colnames(predator_unique_counts) <- c ("method", "predator_species_represented")


p2 <- ggplot (data=predator_unique_counts, aes(y=predator_species_represented, x=method, fill=method)) + geom_bar(stat="identity", color = "black" ) + scale_fill_manual(values=c("#bebada", "#8dd3c7", "#ffaabc", "#ffffb3")) + theme_bw() + ylab ("Predator Species Found") + coord_cartesian(ylim=c(0,50)) + geom_hline(aes (yintercept = length (uniqPredators), linetype = "Total Unique Predator Species"), linetype="dashed", color = "red") + xlab ("Software")

plot (p2)

# For Genus
predator_unique_counts_genus <- cbind.data.frame ( c ("Vsearch", "Ecotag", "Blast", "SAP"),
                                                 c (length (vn_predsGenus[[1]]), length (vn_predsGenus[[2]]), 
                                                    length (vn_predsGenus[[3]]), length (vn_predsGenus[[4]])))
colnames(predator_unique_counts_genus) <- c ("method", "predator_species_represented")


p3 <- ggplot (data=predator_unique_counts_genus, aes(y=predator_species_represented, x=method, fill=method)) + geom_bar(stat="identity", color = "black" ) + scale_fill_manual(values=c("#bebada", "#8dd3c7", "#ffaabc", "#ffffb3")) + theme_bw() + ylab ("Predator Genus Found") + coord_cartesian(ylim=c(0,50)) + geom_hline(aes (yintercept = length (uniqPredatorsGenus), linetype = "Total Unique Predator Genus"), linetype="dashed", color = "red") + xlab ("Software")

plot (p3)
```

## Compare Blast and SAP OTU assignments

SAP should be able to assign wherever Blast did.
However, SAP with default values did not come close. 
In order to experiment with SAP parameters, want to see what OTUs were assigned by Blast (after then id >= .97 filter.)
Of particular interest is the predators, since those assignments are very confident since we know those species should be represented.


```{r goodHitsForSAP, echo = FALSE}

setwd("~/Documents/code/genomics/taxonomic_evalutation/newdata")

# Create comparison table

blast_assignments <- ovt_blast[ovt_blast$BLAST_IDENTITY_NT >= 97,][, c ("OTU_SEQID", "species", "NCBI_TAXID", "BLAST_IDENTITY_NT")]
blast_assignments$IS_PREDATOR <- blast_assignments$species %in% Predators_Samples$SCINAME
sap_assignments <- ovt_sap[, c ("OTU_SEQID", "species", "NCBI_TAXID")]

# What attributes do these SAP assigned OTUs have?
sap_annotations_file <- "classification_SuspiciousCutoff-97.csv"
sap_annotations <- read.csv (file = sap_annotations_file)
colnames (sap_annotations)[1] <- "OTU_SEQID"
# Drop some unneeded columns
sap_annotations$BranchCount <- NULL
sap_annotations$BranchInteriorCount <- NULL
sap_annotations$BranchLeafCount <- NULL
sap_annotations$TreeDepth <- NULL
sap_annotations$HasHomologuesSuspiciousRecords <- NULL


sap_assignments <- merge (x = sap_assignments, y = sap_annotations, by = "OTU_SEQID")

compare_assignments <- merge (x = blast_assignments, y = sap_assignments, by = "OTU_SEQID", all = TRUE)
colnames (compare_assignments)[1:7] = c ("OTU_SEQID", "BLAST_SPECIES", "BLAST_NCBI_TAXID", "BLAST_IDENTITY", "IS_PREDATOR", "SAP_SPECIES", "SAP_NCBI_TAXID")
compare_assignments$IS_MATCH <- compare_assignments$BLAST_SPECIES == compare_assignments$SAP_SPECIES


f <- compare_assignments[!is.na(compare_assignments$BLAST_SPECIES), ]
b <- compare_assignments[ compare_assignments$SAP_SPECIES != "", ]
c <- b[!is.na(b$SAP_SPECIES), ]
total_combined <- length (unique (c (f$OTU_SEQID, c$OTU_SEQID)))


# Summarize comparison table
total_blast <- nrow (blast_assignments)
total_sap <- length (which (sap_assignments$species != ""))
###total_combined <- nrow (compare_assignments)
##len_blast_only <- length (which (is.na (compare_assignments[compare_assignments$SAP_SPECIES != "",]$SAP_SPECIES)))
b <- compare_assignments[!is.na(compare_assignments$BLAST_SPECIES),]
c <- b[b$SAP_SPECIES != "", ]
both <- nrow (c[!is.na(c$SAP_SPECIES),])
len_blast_only <- total_blast - both

len_sap_only <- length (which (is.na (compare_assignments[compare_assignments$SAP_SPECIES != "",]$BLAST_SPECIES)))
# Number of cases where Blast and SAP had same species-level assignment
len_species_match <- length (which (compare_assignments[compare_assignments$SAP_SPECIES != "",]$IS_MATCH == TRUE))
# Number of cases where Blast and SAP both had species-level assignments, but mismatched
len_species_mismatch <- length (which (compare_assignments[compare_assignments$SAP_SPECIES != "",]$IS_MATCH == FALSE))


# Number of cases where SAP assigned to a higher rank
len_nonspecies_sap <- length (which (compare_assignments[compare_assignments$SAP_SPECIES == "",]$SAP_SPECIES == ""))
# Number of cases where SAP had a species-level assignment, but BLAST did not
len_sap_only_species_assignment <- length (which (compare_assignments[is.na (compare_assignments$BLAST_SPECIES), ]$SAP_SPECIES != ""))
# Ensure good case checking

write.csv (compare_assignments, file = "blastSAPcomparison.csv", quote = FALSE, row.names = FALSE)

```

| Intersection | Number of Assigned OTUs |
  | - | -:
  Blast assignments | `r total_blast`
SAP assignments | `r total_sap`
Total OTU assignments | `r total_combined`
OTUs that only Blast assigned | `r len_blast_only`
OTUs that only SAP assigned | `r len_sap_only`
Cases where Blast and SAP assigned same species | `r len_species_match`
Cases where Blast and SAP both assigned species, but mismatched | `r len_species_mismatch`
Cases where SAP did a higher-level assignment | `r len_nonspecies_sap`
Cases where SAP had a species-level assignment and no Blast assignment | `r len_sap_only_species_assignment`


