---
title: "Evaluation of Taxonomic Assignment Packages"
author: "Evan Krell, Chris Bird, Martin French"
date: "November 18, 2016"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Libraries
library ("pracma")  # For string manipulation functions

```


## Load and prepare data

```{r load}

setwd("~/taxonomic_evalutation/data")

# File that maps OTUs to all individual sequences contained
OTUs_seqs_file <- "Simons.combo.nonchimeras.clean.OTU.samples_counts.BEST.csv"
OTUs_seqs_samples <- read.table (file = OTUs_seqs_file, header = FALSE, sep = ',', stringsAsFactors = FALSE)
colnames (OTUs_seqs_samples) <- c ("OTU_SEQID", "QSEQID", "SAMPLE_ID", "COUNT")
OTUs_seqs <- cbind.data.frame (OTUs_seqs_samples$OTU_SEQID, OTUs_seqs_samples$QSEQID)
colnames (OTUs_seqs) <- c ("OTU_SEQID", "QSEQID")

# List of OTUs
OTUs <- OTUs_seqs[!duplicated (OTUs_seqs$OTU_SEQID),]
OTUs <- as.list (OTUs$OTU_SEQID)

# CrittersVSTubes files
cvt_vsearch_file <- "Simons.OTUvsTubes_VsearchXGenbank_with_SCORES.csv"
cvt_ecotag_file <- "Simons.OTUvsTubes_EcotagXGenbank_with_SCORES.csv"
cvt_blast_file <- "Simons.OTUvsTubes_Blast2XGenbank_with_SCORES.fix.csv"
cvt_vsearch <- read.table (file = cvt_vsearch_file, header = TRUE, sep = ',', stringsAsFactors = FALSE)
cvt_ecotag <- read.table (file = cvt_ecotag_file, header = TRUE, sep = ',', stringsAsFactors = FALSE)
cvt_blast <- read.table (file = cvt_blast_file,  header = TRUE, sep = ',', stringsAsFactors = FALSE)
# Adjust VSEARCH Target SEQID so that all use same SEQID scheme
cvt_vsearch$TSEQID <- gsub (x = cvt_vsearch$TSEQID, pattern = "\\|.*$", replacement = "")
# Adjust VSEARCH Identity score to be a decimal
cvt_vsearch$VSEARCH_IDENTITY_VSEARCH_GENBANK <- cvt_vsearch$VSEARCH_IDENTITY_VSEARCH_GENBANK / 100
# Adjust BLAST Target SEQID so that all use same SEQID scheme
cvt_blast$SSEQID <- gsub (x = cvt_blast$SSEQID, pattern = "\\|.*$", replacement = "")
# Adjust VSEARCH Identity score to be a decimal
cvt_blast$BLAST_IDENTITY_BLAST_GENBANK <- cvt_blast$BLAST_IDENTITY_BLAST_GENBANK / 100

# Merge OTU into CVT
ovt_vsearch <- merge (x = OTUs_seqs, y = cvt_vsearch, all = TRUE, by.x = "QSEQID", by.y = "SEQID")
ovt_ecotag <- merge (x = OTUs_seqs, y = cvt_ecotag, all = TRUE, by.x = "QSEQID", by.y = "SEQID")
ovt_blast <- merge (x = OTUs_seqs, y = cvt_blast, all = TRUE, by.x = "QSEQID", by.y = "SEQID")

# Charon files
charon_vsearch_file <- "Simons.combo.vsearch.charon_OTU.csv"
charon_ecotag_file <- "Simons.combo.ecotag.charon_OTU.csv"
charon_blast_file <- "Simons.combo.blast.charon_OTU.csv"
charon_vsearch <- read.table (file = charon_vsearch_file, header = FALSE, sep = ',', stringsAsFactors = FALSE)
charon_ecotag <- read.table (file = charon_ecotag_file, header = FALSE, sep = ',', stringsAsFactors = FALSE)
charon_blast <- read.table (file = charon_blast_file, header = FALSE, sep = ',', stringsAsFactors = FALSE)
colnames (charon_vsearch) <- c ("QUERY_SEQID", "REFERENCE_SEQID", "SCINAME", "NCBI_TAXID", "SAMPLE_ID", "COUNT", "OTU_SEQID")
colnames (charon_ecotag) <- c ("QUERY_SEQID", "REFERENCE_SEQID", "SCINAME", "NCBI_TAXID", "SAMPLE_ID", "COUNT", "OTU_SEQID")
colnames (charon_blast) <- c ("QUERY_SEQID", "REFERENCE_SEQID", "SCINAME", "NCBI_TAXID", "SAMPLE_ID", "COUNT", "OTU_SEQID")
# Adjust VSEARCH Target SEQID so that all use same SEQID scheme
charon_vsearch$REFERENCE_SEQID <- gsub (x = charon_vsearch$REFERENCE_SEQID, pattern = "\\|.*$", replacement = "")
# Adjust BLAST Target SEQID so that all use same SEQID scheme
charon_blast$REFERENCE_SEQID <- gsub (x = charon_blast$REFERENCE_SEQID, pattern = "\\|.*$", replacement = "")

# Compress Charon files to be only OTUs with info
otu_data_vsearch <- charon_vsearch[!duplicated (charon_vsearch$OTU_SEQID),]
otu_data_vsearch$QUERY_SEQID <- NULL
otu_data_vsearch$SAMPLE_ID <- NULL
otu_data_vsearch$COUNT <- NULL

otu_data_ecotag <- charon_ecotag[!duplicated (charon_ecotag$OTU_SEQID),]
otu_data_ecotag$QUERY_SEQID <- NULL
otu_data_ecotag$SAMPLE_ID <- NULL
otu_data_ecotag$COUNT <- NULL

otu_data_blast <- charon_blast[!duplicated (charon_blast$OTU_SEQID),]
otu_data_blast$QUERY_SEQID <- NULL
otu_data_blast$SAMPLE_ID <- NULL
otu_data_blast$COUNT <- NULL

```

## Basic stats

Compute ratio of same to different reference sequences obtained from the methods

```{r compare}
total_OTUs <- length (OTUs)

uniq_vsearch_tseqs <- unique(otu_data_vsearch$REFERENCE_SEQID)
uniq_ecotag_tseqs <- unique(otu_data_ecotag$REFERENCE_SEQID)
uniq_blast_tseqs <- unique(otu_data_blast$REFERENCE_SEQID)

uniq_vsearch_tseqs_len <- length (uniq_vsearch_tseqs)
uniq_ecotag_tseqs_len <- length (uniq_ecotag_tseqs)
uniq_blast_tseqs_len <- length (uniq_blast_tseqs)

intersect_len_EV <- length (intersect (uniq_vsearch_tseqs, uniq_ecotag_tseqs))
interset_len_EB <- length (intersect (uniq_blast_tseqs, uniq_ecotag_tseqs))
intersect_len_VB <- length (intersect (uniq_vsearch_tseqs, uniq_blast_tseqs))
inter <- intersect (uniq_vsearch_tseqs, uniq_ecotag_tseqs)
intersect_len <- length (intersect (inter, uniq_blast_tseqs))
```


| | Vsearch | Ecotag  | Blast
| - |:-:| :-:| -:
| Number of Assigned OTUs | `r nrow (otu_data_vsearch)` | `r nrow (otu_data_ecotag)` | `r nrow (otu_data_blast)`
| Total OTUs | `r total_OTUs` | `r total_OTUs` | `r total_OTUs`
| Percent Assigned OTUs | `r (nrow (otu_data_vsearch)/total_OTUs)*100` | `r (nrow (otu_data_ecotag)/total_OTUs)*100` | `r (nrow (otu_data_blast)/total_OTUs)*100`
| Unique Reference Sequences      | `r uniq_vsearch_tseqs_len` | `r uniq_ecotag_tseqs_len` | ` uniq_blast_tseqs_len`
| Intersection of Unique Reference Sequences        | `r intersect_len`      |   `r intersect_len` | 
| Percent References Sequences Shared | `r (intersect_len/uniq_vsearch_tseqs_len)*100`      |    `r (intersect_len/uniq_ecotag_tseqs_len)*100` | `r (intersect_len/uniq_blast_tseqs_len)*100`


## Investigate phylogenetic levels of taxonomic assingment

```{r phylogeny}
taxonomicRanksOfInterest <- c ("species", "genus", "family", "order", "class", "phylum", "kingdom")

# Vsearch
vsearch_taxonomy <- cvt_vsearch[,taxonomicRanksOfInterest]
vsearch_best_match_scinames <- apply(vsearch_taxonomy, 1, function(x) head(x[x != ""], 1))

vsearch_best_match_scinames <- unlist (lapply(X = vsearch_best_match_scinames, 
                                      function (x) ifelse(length(x) == 0, "", x[1])))
vsearch_best_match_levels <- apply(vsearch_taxonomy, 1, function(x) names(head(x[x != ""], 1)))
vsearch_best_match_levels <- unlist (lapply(X = vsearch_best_match_levels, 
                                      function (x) ifelse(length(x) == 0, "", x[1])))
vsearch_best_match <- cbind.data.frame (cvt_vsearch$SEQID, 
                                        vsearch_best_match_scinames, vsearch_best_match_levels)
colnames (vsearch_best_match) <- c ("QSEQID", "SCINAME", "LEVEL")
vsearch_best_match <- merge (x = OTUs_seqs, y = vsearch_best_match, by = "QSEQID")

# Ecotag
ecotag_taxonomy <- cvt_ecotag[,taxonomicRanksOfInterest]
ecotag_best_match_scinames <- apply(ecotag_taxonomy, 1, function(x) head(x[x != ""], 1))

ecotag_best_match_scinames <- unlist (lapply(X = ecotag_best_match_scinames, 
                                      function (x) ifelse(length(x) == 0, "", x[1])))
ecotag_best_match_levels <- apply(ecotag_taxonomy, 1, function(x) names(head(x[x != ""], 1)))
ecotag_best_match_levels <- unlist (lapply(X = ecotag_best_match_levels, 
                                      function (x) ifelse(length(x) == 0, "", x[1])))
ecotag_best_match <- cbind.data.frame (cvt_ecotag$SEQID, 
                                        ecotag_best_match_scinames, ecotag_best_match_levels)
colnames (ecotag_best_match) <- c ("QSEQID", "SCINAME", "LEVEL")
ecotag_best_match <- merge (x = OTUs_seqs, y = ecotag_best_match, by = "QSEQID")

# Blast
blast_taxonomy <- cvt_blast[,taxonomicRanksOfInterest]
blast_best_match_scinames <- apply(blast_taxonomy, 1, function(x) head(x[x != ""], 1))

blast_best_match_scinames <- unlist (lapply(X = blast_best_match_scinames, 
                                      function (x) ifelse(length(x) == 0, "", x[1])))
blast_best_match_levels <- apply(blast_taxonomy, 1, function(x) names(head(x[x != ""], 1)))
blast_best_match_levels <- unlist (lapply(X = blast_best_match_levels, 
                                      function (x) ifelse(length(x) == 0, "", x[1])))
blast_best_match <- cbind.data.frame (cvt_blast$SEQID, 
                                        blast_best_match_scinames, blast_best_match_levels)
colnames (blast_best_match) <- c ("QSEQID", "SCINAME", "LEVEL")
blast_best_match <- merge (x = OTUs_seqs, y = blast_best_match, by = "QSEQID")




taxon_level_counts_vsearch <- sapply (X = taxonomicRanksOfInterest, 
                                      FUN = function (x) length (which(vsearch_best_match$LEVEL == x)))
taxon_level_counts_ecotag <- sapply (X = taxonomicRanksOfInterest, 
                                     FUN = function (x) length (which(ecotag_best_match$LEVEL == x)))
taxon_level_counts_blast <- sapply (X = taxonomicRanksOfInterest, 
                                     FUN = function (x) length (which(blast_best_match$LEVEL == x)))

empty_taxon_count_vsearch <- length (which (vsearch_best_match$LEVEL == ""))
empty_taxon_count_ecotag <- length (which (ecotag_best_match$LEVEL == ""))
empty_taxon_count_blast <- length (which (blast_best_match$LEVEL == ""))
empty_counts <- c (empty_taxon_count_vsearch, empty_taxon_count_ecotag, empty_taxon_count_blast)

total_count_vsearch <- sum (taxon_level_counts_vsearch) + empty_taxon_count_vsearch
total_count_ecotag <- sum (taxon_level_counts_ecotag) + empty_taxon_count_ecotag
total_count_blast <- sum (taxon_level_counts_blast) + empty_taxon_count_blast
total_counts <- c (total_count_vsearch, total_count_ecotag, total_count_blast)

titles_col <- c ("Vsearch", "Ecotag", "Blast")
counts_table <- rbind.data.frame (taxon_level_counts_vsearch, taxon_level_counts_ecotag, taxon_level_counts_blast)
counts_table <- cbind.data.frame(titles_col, counts_table, empty_counts, total_counts)
colnames (counts_table) <- c ("Source", taxonomicRanksOfInterest, "unassigned", "total")

print (counts_table)

```


## Investigate identity scores

BLAST, Vsearch and Ecotag each return an identity score.

The Vsearch parameters specified a minimum identity score of 0.7.

The Ecotag parameters specified a minumum identity score of ____.

The BLAST parameters specified a minimum identity score of ____.


```{r identity_scores}

## Get each assigned OTU with identity score
otu_identity_vsearch <- cbind.data.frame (ovt_vsearch$OTU_SEQID,
                        ovt_vsearch$VSEARCH_IDENTITY_VSEARCH_GENBANK, ovt_vsearch$TSEQID)
colnames (otu_identity_vsearch) <- c ("OTU_SEQID", "VSEARCH_IDENTITY_GENBANK", "VSEARCH_REF_SEQID")
otu_identity_vsearch <- otu_identity_vsearch[!duplicated (otu_identity_vsearch$OTU_SEQID),]

otu_identity_ecotag <- cbind.data.frame (ovt_ecotag$OTU_SEQID, 
                        ovt_ecotag$ECOTAG_IDENTITY_GENBANK, ovt_ecotag$TSEQID)
colnames (otu_identity_ecotag) <- c ("OTU_SEQID", "ECOTAG_IDENTITY_GENBANK", "ECOTAG_REF_SEQID")
otu_identity_ecotag <- otu_identity_ecotag[!duplicated (otu_identity_ecotag$OTU_SEQID),]
colnames (otu_identity_ecotag) <- c ("OTU_SEQID", "ECOTAG_IDENTITY_GENBANK", "ECOTAG_REF_SEQID")

otu_identity_blast <- cbind.data.frame (ovt_blast$OTU_SEQID,
                        ovt_blast$BLAST_IDENTITY_BLAST_GENBANK, ovt_blast$SSEQID)
colnames (otu_identity_blast) <- c ("OTU_SEQID", "BLAST_IDENTITY_GENBANK", "BLAST_REF_SEQID")
otu_identity_blast <- otu_identity_blast[!duplicated (otu_identity_blast$OTU_SEQID),]

# Together such that each OTU has all identity scores
otu_identity_EV <- merge (x = otu_identity_vsearch, y = otu_identity_ecotag, by = "OTU_SEQID")
otu_identity_EVB <- merge (x = otu_identity_EV, y = otu_identity_blast,  by = "OTU_SEQID")
```

Plot the identity scores for each method separately. 

```{r plot_identity_scores_separate}
par(mfrow=c(1,3))
plot (x = rownames(otu_identity_EVB), y = otu_identity_EVB$VSEARCH_IDENTITY_GENBANK, 
      ylim = c(0.5,1), col = "blue", pch = '.',
      ylab = "Vsearch Identity Score", xlab = "ith OTU")
plot (x = rownames(otu_identity_EVB), y = otu_identity_EVB$ECOTAG_IDENTITY_GENBANK, 
      ylim = c(0.5, 1), col = "red", pch = '.',
      ylab = "Ecotag Identity Score", xlab = "ith OTU")
plot (x = rownames(otu_identity_EVB), y = otu_identity_EVB$BLAST_IDENTITY_GENBANK, 
      ylim = c(0.5, 1), col = "green", pch = '.',
      ylab = "Blast Identity Score", xlab = "ith OTU")
```

Plot the identity scores for each method together.

```{r plot_identity_scores_together}
plot (x = rownames(otu_identity_EVB), y = otu_identity_EVB$VSEARCH_IDENTITY_GENBANK, col = "blue", pch = '.')
points (x = rownames(otu_identity_EVB), y = otu_identity_EVB$ECOTAG_IDENTITY_GENBANK, col = "red", pch = '.')
points (x = rownames(otu_identity_EVB), y = otu_identity_EVB$BLAST_IDENTITY_GENBANK, col = "green", pch = '.')

```


In order to better compare, select OTUs that assigned to the same reference sequences with different methods. 

```{r }
otu_EV_same_refseq <- otu_identity_EV[as.character (otu_identity_EV$VSEARCH_REF_SEQID) == as.character (otu_identity_EV$ECOTAG_REF_SEQID),]
otu_EV_same_refseq <- otu_EV_same_refseq[complete.cases(otu_EV_same_refseq),]
num_same_ref_seq_EV <- nrow (otu_EV_same_refseq)

otu_VB_same_refseq <- otu_identity_EVB[ as.character (otu_identity_EVB$VSEARCH_REF_SEQID) == as.character (otu_identity_EVB$BLAST_REF_SEQID),]
otu_VB_same_refseq <- otu_VB_same_refseq[complete.cases(otu_VB_same_refseq),]
num_same_ref_seq_VB <- nrow (otu_VB_same_refseq)

otu_EB_same_refseq <- otu_identity_EVB[ as.character (otu_identity_EVB$ECOTAG_REF_SEQID) == as.character (otu_identity_EVB$BLAST_REF_SEQID),]
otu_EB_same_refseq <- otu_EB_same_refseq[complete.cases(otu_EB_same_refseq),]
num_same_ref_seq_EB <- nrow (otu_EB_same_refseq)

```

Vsearch and Ecotag share `r num_same_ref_seq_EV` assignments to the same reference sequence.

Vsearch and Blast share `r num_same_ref_seq_VB` assignments to the same reference sequence.

```{r}

identity_diffs_EV <- 
  abs (otu_EV_same_refseq$VSEARCH_IDENTITY_GENBANK - otu_EV_same_refseq$ECOTAG_IDENTITY_GENBANK)

plot (x = seq (1:nrow (otu_EV_same_refseq)), y = identity_diffs_EV, col = "purple", pch = '*',
      ylab = "Vsearch & Ecotag Identity Score Difference", xlab = "ith Reference Sequence")

identity_diffs_VB <- 
  abs (otu_VB_same_refseq$VSEARCH_IDENTITY_GENBANK - otu_VB_same_refseq$BLAST_IDENTITY_GENBANK)

plot (x = seq (1:nrow (otu_VB_same_refseq)), y = identity_diffs_VB, col = "orange", pch = '*',
      ylab = "Vsearch & Blast Identity Score Difference", xlab = "ith Reference Sequence")

```



## Investigate Taxonomy


```{r taxonomy}

# Count number of unique ranks per phylogenetic level

uniqRanks <- cbind ( c ("Vsearch", "Ecotag", "Blast"), rbind (
apply(vsearch_taxonomy, MARGIN = 2, FUN = function (x) length (unique(x))),
apply(ecotag_taxonomy, MARGIN = 2, FUN = function (x) length (unique(x))),
apply(blast_taxonomy, MARGIN = 2, FUN = function (x) length (unique(x))) ) )
colnames (uniqRanks)[1] <- "Source"
print (uniqRanks)

intersect_taxonomy <- function (A, B){
  intersect_lens <- c ()
  for (i in taxonomicRanksOfInterest){
    intersect_lens <- c (intersect_lens, ( length ( intersect ( unlist (unique (A[i])), unlist (unique (B[i])) ) )) )
  }
  intersect_lens
}

intersect_taxonomy_three <- function (A, B, C){
  intersect_lens <- c ()
  for (i in taxonomicRanksOfInterest){
    inter_AB <- intersect (unlist (unique (A[i])), unlist (unique (B[i])) )
    intersect_lens <- c (intersect_lens, ( length ( intersect ( inter_AB, unlist (unique (C[i])) ) )) )
  }
  intersect_lens
}

intersect_counts_EV <- intersect_taxonomy(A = vsearch_taxonomy, B = ecotag_taxonomy)
intersect_counts_EB <- intersect_taxonomy(A = ecotag_taxonomy, B = blast_taxonomy)
intersect_counts_BV <- intersect_taxonomy(A = vsearch_taxonomy, B = blast_taxonomy)
intersect_counts_EBV <- intersect_taxonomy_three(A = vsearch_taxonomy, B = ecotag_taxonomy, C = blast_taxonomy)
intersect_counts_table <- cbind.data.frame (intersect_counts_EV, intersect_counts_EB, intersect_counts_BV, intersect_counts_EBV)
colnames (intersect_counts_table) <- c ("Vsearch, Ecotag", "Ecotag, Blast", "Vsearch, Blast", "Vsearch, Ecotag, Blast")

print (intersect_counts_table)


uniqRanks <- cbind ( c ("Vsearch", "Ecotag", "Blast"), rbind (
apply(vsearch_taxonomy, MARGIN = 2, FUN = function (x) length (unique(x))),
apply(ecotag_taxonomy, MARGIN = 2, FUN = function (x) length (unique(x))),
apply(blast_taxonomy, MARGIN = 2, FUN = function (x) length (unique(x))) ) )
colnames (uniqRanks)[1] <- "Source"

print (uniqRanks)

# Get counts of specific unique ranks within a level
vsearch_taxonomy_uniq <- apply(vsearch_taxonomy, MARGIN = 2, FUN = function (x) unique(x))
ecotag_taxonomy_uniq <- apply(ecotag_taxonomy, MARGIN = 2, FUN = function (x) unique(x))
blast_taxonomy_uniq <- apply(blast_taxonomy, MARGIN = 2, FUN = function (x) unique(x))

# Vsearch class counts
vsearch_class_counts <- sapply (vsearch_taxonomy_uniq$class, FUN = function (x) length (which(vsearch_taxonomy$class == x)))
vsearch_phylum_counts <- sapply (vsearch_taxonomy_uniq$phylum, FUN = function (x) length (which(vsearch_taxonomy$phylum == x)))
vsearch_kingdom_counts <- sapply (vsearch_taxonomy_uniq$kingdom, FUN = function (x) length (which(vsearch_taxonomy$kingdom == x)))

# Ecotag class counts
ecotag_class_counts <- sapply (ecotag_taxonomy_uniq$class, FUN = function (x) length (which(ecotag_taxonomy$class == x)))
ecotag_phylum_counts <- sapply (ecotag_taxonomy_uniq$phylum, FUN = function (x) length (which(ecotag_taxonomy$phylum == x)))
ecotag_kingdom_counts <- sapply (ecotag_taxonomy_uniq$kingdom, FUN = function (x) length (which(ecotag_taxonomy$kingdom == x)))

# Blast class counts
blast_class_counts <- sapply (blast_taxonomy_uniq$class, FUN = function (x) length (which(blast_taxonomy$class == x)))
blast_phylum_counts <- sapply (blast_taxonomy_uniq$phylum, FUN = function (x) length (which(blast_taxonomy$phylum == x)))
blast_kingdom_counts <- sapply (blast_taxonomy_uniq$kingdom, FUN = function (x) length (which(blast_taxonomy$kingdom == x)))

print (vsearch_class_counts)
print (ecotag_class_counts)
print (blast_class_counts)

print (vsearch_phylum_counts)
print (ecotag_phylum_counts)
print (blast_phylum_counts)

print (vsearch_kingdom_counts)
print (ecotag_kingdom_counts)
print (blast_kingdom_counts)
```




